% novelty-theory-lean — paper skeleton
% Preamble aligned with ugp-lean / NEMS suite conventions + Representational_Incompleteness \leanref.
% Build (from ./paper):  latexmk -pdf novelty_theory_lean_fixed_generation.tex
%   or:  pdflatex → bibtex → pdflatex ×2

\documentclass[11pt,a4paper]{article}

% ── Packages (ugp-lean-style backbone) ───────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{microtype}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{cleveref}
\usepackage{needspace}
\usepackage[most]{tcolorbox}

\setlength{\hfuzz}{3pt}

% ── Line breaks in long Lean identifiers (Representational_Incompleteness.tex) ──
\usepackage{expl3}
\usepackage{xparse}
\ExplSyntaxOn
\tl_new:N \l__ntl_leanref_tl
\NewDocumentCommand \leanref { m }
  {
    \tl_set:Nn \l__ntl_leanref_tl { #1 }
    \regex_replace_all:nnN { \c{_} } { \c{_} \c{allowbreak} } \l__ntl_leanref_tl
    \regex_replace_all:nnN { \. } { . \c{allowbreak} } \l__ntl_leanref_tl
    \regex_replace_all:nnN { ([a-z])([A-Z]) } { \1 \c{allowbreak} \2 } \l__ntl_leanref_tl
    \regex_replace_all:nnN { ([0-9])([A-Z]) } { \1 \c{allowbreak} \2 } \l__ntl_leanref_tl
    \texttt{\small \tl_use:N \l__ntl_leanref_tl }
  }
\ExplSyntaxOff

\newenvironment{keybox}[1]{%
  \Needspace*{8\baselineskip}%
  \begin{tcolorbox}[
    enhanced, breakable=false,
    colback=blue!5, colframe=blue!45!black, coltitle=black,
    colbacktitle=blue!14, boxrule=0.85pt, arc=2.5pt,
    left=9pt, right=9pt, top=6pt, bottom=6pt,
    titlerule=0.55pt, fonttitle=\bfseries, title={#1},
    before upper=\raggedright,
    before skip=0.9\baselineskip, after skip=0.9\baselineskip
  ]%
}{%
  \end{tcolorbox}%
}

% ── Lean listing style (ugp-lean) ─────────────────────────────
\definecolor{leanblue}{RGB}{0,0,180}
\definecolor{leangreen}{RGB}{0,128,0}
\definecolor{leangray}{RGB}{128,128,128}
\lstdefinelanguage{Lean4}{
  keywords={theorem,def,lemma,example,structure,namespace,end,import,
            open,section,variable,where,by,exact,have,let,show,calc,
            intro,apply,rw,simp,omega,ring,linarith,native_decide,
            unfold,obtain,rcases,use,constructor,refine,induction,with},
  keywordstyle=\color{leanblue}\bfseries,
  commentstyle=\color{leangray}\itshape,
  stringstyle=\color{leangreen},
  morecomment=[l]{--},
  morecomment=[n]{/-}{-/},
  sensitive=true,
  basicstyle=\ttfamily\small,
  breaklines=true,
  columns=flexible,
  literate={→}{$\to$}1 {←}{$\leftarrow$}1 {∀}{$\forall$}1 {∃}{$\exists$}1
           {∧}{$\wedge$}1 {∨}{$\vee$}1 {¬}{$\neg$}1 {≤}{$\le$}1
           {≥}{$\ge$}1 {≠}{$\ne$}1 {∈}{$\in$}1 {ℕ}{$\mathbb{N}$}1
}
\lstset{language=Lean4}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\lean}{\textsc{Lean}\,4}

\title{\textbf{novelty-theory-lean}:\\
  Fixed Generation Without Fixed Explanatory Closure\\
  {\large Machine-Checked Generative Systems, Stratified Provability, and Crown Packaging}}
\hypersetup{
  pdftitle={novelty-theory-lean: Fixed generation without fixed explanatory closure}
}
\author{Nova Spivack\\
  \small\url{https://github.com/novaspivack/novelty-theory-lean}}
\date{Draft skeleton --- \today}

\begin{document}
\maketitle

\begin{abstract}
\raggedright
\noindent\textbf{First sentence (existential flagship).}
There exists a concrete deterministic generative system on~$\mathbb{N}$
(the library's canonical counter) and a packaged sentence universe such
that a family of \emph{structurally certified} sentences exhibits strict
proof-theoretic ascent under an explicit depth calculus, while generator
truth on the same counter coexists with this phenomenon in a single
bundled statement.

\medskip
\noindent\textbf{Second sentence (class-transfer strengthening).}
Beyond a single model artifact, the library constructs an admissible
class of aligned $\mathbb{N}\to(\mathrm{Bool}\times\mathbb{N})$
generators for which the same enum-crown gap template transfers while
the Bool observation disagrees everywhere---so the phenomenon is not tied
to one carrier tag by construction.

\medskip
\noindent\textbf{Third sentence (non-finality and terminality defeat).}
The generalized summit also packages scoped non-finality of an
explanatory regime diagram (as formalized) and a formal future-defeat
property for bounded terminality-style claims on the canonical counter,
alongside organization obstructions stated at bounded multiscale height.

\medskip
\noindent\textbf{Trust boundary.}
No literal universal over all raw admissible-interface rows, and no
universal infinite paradigm shift for every singleton-within regime
family, is claimed or available: such universal glosses are \emph{false}
under the formalized predicates and are explicitly refuted by
obstruction lemmas in the artifact. All positive theorems are scoped to
the definitions exported from the \lean{} library \cite{noveltyTheoryLean2026}.
The formalization is zero-\texttt{sorry} on the packaged library path
and builds against Mathlib~\cite{mathlib4}, toolchain in the repository pin~\cite{deMoura2021}.
\end{abstract}

\tableofcontents
\bigskip

\section{Premises, trust boundary, and what can still be attacked}\label{sec:trust}

\begin{keybox}{Where disagreement can (and cannot) live}
\textbf{Formal spine.}
All mathematical claims in this paper, when stated as theorems, refer to
definitions and proofs in the publicly buildable \lean{} library
\cite{noveltyTheoryLean2026}. Section~\ref{sec:appendix-map} lists a
theorem-to-prose map; Section~\ref{sec:appendix-dag} shows the informal
layer graph.

\medskip
\textbf{Premise coordinates (for external readers).}
\begin{itemize}[leftmargin=2em]
\item \textbf{Proof theory vs.\ metaphysics.} ``Strict $\vdash$ / $\not\vdash$'' is
  internal to the Model~C depth calculus formalized in the library, not a claim
  about all possible logics or all physical universes.
\item \textbf{Generation $\neq$ explanation $\neq$ provability.} Section~\ref{sec:three-levels}
  keeps these layers explicit; collapsing them in prose would overstate the formal result.
\item \textbf{Scoped admissibility.} Class-transfer results quantify over the
  \emph{constructed} aligned carriers used in the proofs---not an
  unqualified ``all mathematical dynamical systems'' statement.
\item \textbf{Blocked literal universals.} Two universal templates are
  \emph{refuted} in-library (see Section~\ref{sec:blocked-universals}): raw
  interface diagonals without further scoping, and infinite paradigm chains for every
  singleton-within regime family.
\end{itemize}
\end{keybox}

\noindent\textbf{Keywords:}
interactive theorem proving, \lean, formalization, generative systems,
stratified provability, explanatory regimes, anti-collapse, crown packaging

\section{Introduction (outline)}\label{sec:intro}

\paragraph{Problem coordinate.}
\emph{[Draft: fixed generation without fixed explanatory closure; observation/simulation need not explanatory reduction.]}

\paragraph{Contribution (frozen flagship).}
The primary named package is
\texttt{NoveltyTheory.Summits.GeneralizedFinalCrownPackage}; supporting roles for earlier summits and \texttt{FinalCrownPackage} without coequal headline status.

\paragraph{Roadmap.}
Section~\ref{sec:blocked-universals}: boundaries first.
Section~\ref{sec:three-levels}: generation / explanation / provability.
Section~\ref{sec:anti-collapse}: why the crown is not a mention-bound artifact.
Section~\ref{sec:comparison}: genotype/phenotype, computational irreducibility, emergence rhetoric.
Section~\ref{sec:formal-overview}: informal tour of the calculus.
Section~\ref{sec:discussion}: limitations and future frontier (\emph{richer phase-at-sentence modalities, broader carriers}).

\section{Blocked universal templates (write boundaries before the intro polish)}\label{sec:blocked-universals}

\paragraph{Raw admissible interface.}
\emph{[Draft: cite obstruction lemmas: trivial interfaces can make diagonal rows exhaustive; honest packaging uses signature-scoped or bundled admissibility.]}

\paragraph{Singleton-within regime families.}
\emph{[Draft: empty regime / obstruction to universal infinite paradigm chain.]}

\section{Three levels of novelty (do not let readers blur these)}\label{sec:three-levels}

\paragraph{Generation (dynamics / trace level).}
\emph{[Draft.]}

\paragraph{Explanation (regime / phase / organizational diagrams).}
\emph{[Draft.]}

\paragraph{Provability and expressibility (sentence height / Model~C).}
\emph{[Draft.]}

\section{Anti-collapse: what the crown is demonstrably not}\label{sec:anti-collapse}

\paragraph{Not only \texttt{geOutput}.}
\emph{[Draft: structural witness, enum crown, structural v2 / hist bundles.]}

\paragraph{Not only trace-ridge atoms.}
\emph{[Draft.]}

\paragraph{Not only \texttt{mentionBound} stratification.}
\emph{[Draft: paired sentences at same stratum, distinct formulas.]}

\paragraph{Not a single-model artifact (Bool-tagged class).}
\emph{[Draft: link
\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_not\_model\_artifact}.]}

\section{Comparison with nearby narratives}\label{sec:comparison}

\paragraph{Genotype / phenotype and simple adequacy readings.}
\emph{[Draft keyed to observation $\not\Rightarrow$ explanatory reduction lemmas.]}

\paragraph{Computational irreducibility \cite{Wolfram2002} and Rule~110 \cite{Cook2004}.}
\emph{[Draft: conceptual kinship vs.\ formal claim mismatch---this paper is proof-theoretic and regime-organizational, not a cellular automaton undecidability export.]}

\paragraph{``Emergence'' rhetoric.}
\emph{[Draft: only what \lean{} exports.]}

\section{Formal setting (informal)}\label{sec:formal-overview}

\paragraph{Generative systems and the canonical counter.}
\emph{[Draft: states, step, trace, output.]}

\paragraph{Sentences, semantics, and \texttt{ProvesAt}.}
\emph{[Draft: Model~C bridge; depth as a resource.]}

\paragraph{Regimes, phases, organization.}
\emph{[Draft: \texttt{explains}, bounded organization predicates.]}

\section{Flagship generalized package (prose target)}\label{sec:flagship}

\emph{[Draft: restate the six \texttt{generalized\_final\_crown\_*} theorems informally; point to Appendix~\ref{sec:appendix-map} for exact names and boundaries.]}

\section{What this paper does \emph{not} prove (one-page discipline)}\label{sec:not-prove}

\begin{itemize}[leftmargin=2em]
\item Not: every deterministic system exhibits the strict-gap pattern.
\item Not: every explanatory formalization fails globally.
\item Not: every organization must be infinite (only bounded packaged obstructions as stated).
\item Not: metaphysical anti-determinism or backward causation.
\item Not: a universal diagonal over all raw interface rows, or a universal infinite paradigm tower.
\end{itemize}

\section{Discussion and future frontier}\label{sec:discussion}

\emph{[Draft: richer phase-at-sentence without \texttt{Set}\,$\mathbb{N}$ encoding; extra modalities; additional carriers---explicitly ``new program'' beyond the frozen tranche.]}

\section{Conclusion}\label{sec:conclusion}

\emph{[Draft: freeze claim surface; point to repository build instructions.]}

\paragraph{Reproducibility.}
From the repository root, \texttt{lake build} compiles \texttt{NoveltyTheory}; the paper bundle lives under \path{paper/}.

% ═══════════════════════════════════════════════════════════════
\appendix

\section{Theorem-to-prose map}\label{sec:appendix-map}
\input{figures/theorem_to_prose_table}

\section{Human-readable dependency layers}\label{sec:appendix-dag}
\input{figures/layer_dependency}

\section{Appendix-style theorem inventory (dependency-focused)}\label{sec:appendix-inventory}

\paragraph{Summit stacks (supporting layers).}
\emph{[Draft: SummitPackages / OrganizationFrontier / FinalCrownPackage as supporting narrative, not coequal headlines.]}

\paragraph{Crown completion layer.}
\emph{[Draft: enum / poly / org V2 / retro v2 lemmas with representative \leanref{} anchors.]}

\paragraph{Obstruction lemmas preventing universal glosses.}
\emph{[Draft: name representative counterexample interfaces / empty regimes without internal repo paths in external PDF---use \leanref{} only.]}

\bibliographystyle{plain}
\bibliography{refs}

\end{document}
