\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{microtype}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{seqsplit}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{cleveref}
\usepackage{needspace}
\usepackage[most]{tcolorbox}
\usepackage{expl3}
\usepackage{xparse}

\setlength{\hfuzz}{3pt}

\ExplSyntaxOn
\tl_new:N \l__ntl_leanref_tl
\NewDocumentCommand \leanref { m }
  {
    \tl_set:Nn \l__ntl_leanref_tl { #1 }
    \regex_replace_all:nnN { \c{_} } { \c{_} \c{allowbreak} } \l__ntl_leanref_tl
    \regex_replace_all:nnN { \. } { . \c{allowbreak} } \l__ntl_leanref_tl
    \regex_replace_all:nnN { ([a-z])([A-Z]) } { \1 \c{allowbreak} \2 } \l__ntl_leanref_tl
    \regex_replace_all:nnN { ([0-9])([A-Z]) } { \1 \c{allowbreak} \2 } \l__ntl_leanref_tl
    \texttt{\small \tl_use:N \l__ntl_leanref_tl }
  }
\ExplSyntaxOff

\newenvironment{keybox}[1]{%
  \Needspace*{8\baselineskip}%
  \begin{tcolorbox}[
    enhanced, breakable=false,
    colback=blue!5, colframe=blue!45!black, coltitle=black,
    colbacktitle=blue!14, boxrule=0.85pt, arc=2.5pt,
    left=9pt, right=9pt, top=6pt, bottom=6pt,
    titlerule=0.55pt, fonttitle=\bfseries, title={#1},
    before upper=\raggedright,
    before skip=0.9\baselineskip, after skip=0.9\baselineskip
  ]%
}{%
  \end{tcolorbox}%
}

% Compact, skimmable proof / witness notes (lighter than keybox).
\newenvironment{proofsketch}{%
  \par\smallskip
  \begin{tcolorbox}[
    enhanced, breakable,
    colback=blue!4, colframe=blue!28, boxrule=0.55pt, arc=2pt,
    left=8pt, right=8pt, top=5pt, bottom=5pt,
    before skip=0.35\baselineskip, after skip=0.45\baselineskip,
    fontupper=\small
  ]
  \noindent\textbf{Proof sketch.}\quad\ignorespaces
}{\end{tcolorbox}}

\newenvironment{witnessmeta}{%
  \par\smallskip
  \begin{tcolorbox}[
    enhanced, breakable=false,
    colback=gray!7, colframe=gray!55, boxrule=0.45pt, arc=2pt,
    left=8pt, right=8pt, top=5pt, bottom=5pt,
    before skip=0.25\baselineskip, after skip=0.4\baselineskip,
    fontupper=\small
  ]
}{\end{tcolorbox}}

\definecolor{leanblue}{RGB}{0,0,180}
\definecolor{leangreen}{RGB}{0,128,0}
\definecolor{leangray}{RGB}{128,128,128}
\lstdefinelanguage{Lean4}{
  keywords={theorem,def,lemma,example,structure,namespace,end,import,
            open,section,variable,where,by,exact,have,let,show,calc,
            intro,apply,rw,simp,omega,ring,linarith,native_decide,
            unfold,obtain,rcases,use,constructor,refine,induction,with},
  keywordstyle=\color{leanblue}\bfseries,
  commentstyle=\color{leangray}\itshape,
  stringstyle=\color{leangreen},
  morecomment=[l]{--},
  morecomment=[n]{/-}{-/},
  sensitive=true,
  basicstyle=\ttfamily\small,
  breaklines=true,
  columns=flexible,
  literate={→}{$\to$}1 {←}{$\leftarrow$}1 {∀}{$\forall$}1 {∃}{$\exists$}1
           {∧}{$\wedge$}1 {∨}{$\vee$}1 {¬}{$\neg$}1 {≤}{$\le$}1
           {≥}{$\ge$}1 {≠}{$\ne$}1 {∈}{$\in$}1 {ℕ}{$\mathbb{N}$}1
}
\lstset{language=Lean4}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\lean}{\textsc{Lean}\,4}
\newcommand{\NT}{\mathsf{NT}}
\newcommand{\Gen}{\mathsf{Gen}}
\newcommand{\Reg}{\mathsf{Reg}}
\newcommand{\Phase}{\mathsf{Phase}}
\newcommand{\Expl}{\mathsf{Expl}}
\newcommand{\Adeq}{\mathsf{Adeq}}
\newcommand{\Reduce}{\mathsf{Reduce}}
\newcommand{\ConservativeOver}{\mathsf{ConservativeOver}}
\newcommand{\GeneratedBy}{\mathsf{GeneratedBy}}
\newcommand{\ParadigmShift}{\mathsf{ParadigmShift}}
\newcommand{\ProvesAt}{\mathsf{ProvesAt}}
\newcommand{\HoldsAt}{\mathsf{HoldsAt}}
\newcommand{\ExpressibleAt}{\mathsf{ExpressibleAt}}
\newcommand{\FutureDefeat}{\mathsf{FutureDefeat}}
\newcommand{\Terminal}{\mathsf{Terminal}}
\newcommand{\Structural}{\mathsf{Structural}}
\newcommand{\Admissible}{\mathsf{Adm}}
% Breakable typewriter tokens (e.g.\ long identifiers in narrow table columns)
\newcommand{\ttwrap}[1]{{\ttfamily\seqsplit{#1}}}

\title{\textbf{Self-Transcending Generators:}\\
Fixed Causal Laws Without Final Explanatory Closure}
\author{Nova Spivack}
\date{}

\begin{document}
\maketitle

\begin{abstract}
\raggedright
\noindent
\textbf{Main claim.} A common tacit assumption is that if a phenomenon is generated by fixed fundamental laws, then there is (in principle) a fixed final explanatory vocabulary in which everything important about that phenomenon can be said. That assumption is \emph{false} in a precise, constructive sense: there are finitely specified lawful generators whose entire phase futures lie in a single causal trace, yet no fixed admissible explanatory interface closes the tower.

\smallskip
\noindent
\textbf{What is proved.} The paper presents a coordinated theorem family whose spine is organized around four main thematic peaks: (i)~\emph{causal completeness without explanatory finality}---every phase is generated forward, while each fixed admissible reducer fails somewhere; (ii)~\emph{infinite conservative paradigm succession}---later regimes can preserve what earlier ones secured and still force irreducibly new structure; (iii)~\emph{upward explanatory necessity}---later regimes can be \emph{indispensable} for stating or proving structural truths about the generator itself (the ``crown'' inversion); and (iv)~\emph{generalized transfer}---the pattern is not a coding accident of one model: it survives disciplined transport across a nontrivial admissible class, including output-enum strict gaps on every aligned natural carrier, obstructions for bounded natural organization against certifying the full standard future, and packaged anti-closure on explicit natural numeric predicates. A further machine-checked layer records \emph{sharp boundaries}: packaged paradigm succession is equivalent to a weak step \emph{and} history conservativity (so dropping conservativity is not a harmless abbreviation); on numeric observation type, unbounded observable recurrence is equivalent to failing a uniform trace bound; and unbounded recurrence is \emph{not} a consequence of lawful generation alone (explicit counterexamples exist). None of the positive results require indeterminism, backward causation, or blanket universals over arbitrary interfaces; the quantifiers in the formal statements mark exactly what is---and is not---asserted.

\smallskip
\noindent
\textbf{Formal verification.} The theorem statements discussed here are machine-checked in a companion Lean~4 development. The scope of every formal claim is fixed by the cited declarations and their explicit hypotheses.

\smallskip
\noindent
\textbf{Philosophical payoff.} Causal generation from a fixed law and explanatory closure are shown to diverge: a world can be fully lawful and fully generated in the forward sense while remaining \emph{in principle} resistant to finitely closed explanation relative to the relevant admissible class. That is \textbf{lawful self-transcendence}: upward-generated structure that is not explanation-terminal from below.

\smallskip
\noindent
\textbf{Trust boundary.} Machine-checked claims are exactly those in the cited \lean{} declarations (especially \Cref{tab:crown-lean}, \Cref{tab:strengthened-lean}, \Cref{tab:theorem-inventory}); they do not assert ``every deterministic system'' or a universal diagonal over arbitrary raw interfaces. \Cref{sec:premises-novelty-crown} fixes \emph{where disagreement can and cannot live}; \Cref{sec:boundaries} restates scope. Interpretive philosophy in \Cref{sec:philosophical} is labeled as such and does not strengthen quantifiers.
\end{abstract}

\noindent\textbf{Keywords:} lawful generation, explanatory closure, paradigm shift, proof-theoretic ascent, upward explanatory necessity, Lean formalization, admissible reducers, self-transcending generators.

\noindent\textbf{MSC 2020:} 03B70, 03F30, 03A05, 68T27.

\section{Premises, disagreement coordinates, and trust boundary}\label{sec:premises-novelty-crown}
\begin{keybox}{Where disagreement can (and cannot) live}
\textbf{Formal spine.}
The generalized final crown package (six coordinated clauses), lower summits, obstructions, and naturality-aware transport are proved in \texttt{NoveltyTheory/} as cited; for a full machine-checked audit trail see \Cref{tab:crown-lean}--\Cref{tab:theorem-inventory} and \Cref{sec:witnesses-proofs}.

\medskip
\textbf{Premise coordinates (for external readers).}
{\raggedright
\begin{itemize}[leftmargin=2em]
\item \textbf{Global metaphysics vs.\ formal quantifiers.} Theorems refute fixed final \emph{admissible} explanatory closure in the constructed models and under the predicates stated (including natural-class hypotheses where used). A blanket claim that ``no final vocabulary is possible for all phenomena everywhere'' is \emph{not} what is discharged.
  \emph{Logical effect:} Map polemics to \Cref{tab:crown-lean}, \Cref{tab:strengthened-lean}, and \Cref{sec:boundaries}.

\item \textbf{Causal completeness vs.\ explanatory finality.} The paper denies that lawful generation forces explanatory closure in the relevant sense; it does \emph{not} deny forward causation or that phases lie in the generator's trace.
  \emph{Logical effect:} Upward necessity (\Cref{sec:upward-necessity}) concerns proof- and regime-relative access, not backward causation (\Cref{sec:order-divergence}).

\item \textbf{Coding trick vs.\ class transfer.} A narrow ``only one model'' objection is met, under explicit hypotheses, by class transfer and the naturality-aware block.
  \emph{Logical effect:} Compare the objection's quantifiers to \Cref{sec:generalized-crown} and \Cref{tab:strengthened-lean}.

\item \textbf{Universal inevitability vs.\ cited predicates.} ``Every lawful generator outruns every numeric bound'' and ``every weak paradigm step packages a full shift'' are \emph{false} under unrestricted quantifiers; what is proved are structured witnesses and, where relevant, categorical equivalences and obstructions at the stated layer (\Cref{sec:boundaries}, \Cref{sec:witnesses-proofs}).
  \emph{Logical effect:} Map sweeping glosses to \Cref{tab:theorem-inventory} and the obstruction lemmas named there.
\end{itemize}
\par}
\end{keybox}

\tableofcontents

\section{Introduction}\label{sec:intro}

\begin{quote}
If a phenomenon is generated by fixed fundamental laws, then in principle there exists a fixed final explanatory language in which everything important about that phenomenon can be said.
\end{quote}

That creed is one of the deepest unexamined assumptions in modern thought. The theory developed here overturns it by exhibiting a sharp mathematical profile and proving its stability under transfer.

The thesis is not that laws fail, that causation fails, or that later phases are ungenerated. It is sharper: a system can be causally complete without being explanatorily final. A fixed law can produce every phase that appears and still defeat every would-be final reducer drawn from a nontrivial admissible class.

\smallskip
\noindent
\textbf{Formal backbone.} The argument is paired with a machine-checked Lean~4 development whose cited declarations fix the exact formal scope of the claims proved here.

\begin{center}
\fbox{\begin{minipage}{0.92\linewidth}
\textbf{Honesty (scope).} Not every deterministic system need realize strict depth-indexed ascent or full crown packaging. Where the formal predicates are not assumed, the theorems are silent. Naive universal diagonals over raw admissible interfaces are explicitly blocked, not proved (\Cref{sec:boundaries}, obstruction lemmas in the library).
\end{minipage}}
\end{center}

\begin{center}
\fbox{\begin{minipage}{0.92\linewidth}
\textbf{Headline statements (preview)}\\[0.35em]
\textbf{(1) Non-finality.} Fixed lawful generation does not force a fixed final admissible explanatory stance for the whole tower (\Cref{sec:first-summit}, \Cref{sec:no-final-internal}).\\
\textbf{(2) Generalized crown.} Iterated structural ascent, class transfer, future defeat of terminality, generator-truth alignment, and non-single-model stability are packaged as one backbone (\Cref{sec:generalized-crown}).\\
\textbf{(3) Crown inversion.} Later regimes can be necessary for structural truths about the generator itself (\Cref{sec:upward-necessity}).
\end{minipage}}
\end{center}

\medskip
\noindent
The results ascend through the following summits (expository order):
\begin{enumerate}[label=\textbf{Summit \Roman*.},leftmargin=*]
\item \textbf{Triviality guard and schema.} What would be cheap, and what conjunction is actually targeted (\Cref{sec:nottrivial}, \Cref{sec:causal-vs-explanatory}).
\item \textbf{Object layer.} Self-transcending generators and conservative paradigm shift as formal objects (\Cref{sec:self-transcending}, \Cref{sec:paradigm-shift}).
\item \textbf{Fixed generation without explanatory terminality.} Every phase on the trace, yet every fixed admissible reducer fails somewhere (\Cref{sec:first-summit}).
\item \textbf{No final internal theory.} Finality for the generated future is impossible in the relevant class (\Cref{sec:no-final-internal}).
\item \textbf{Generalized final crown.} Class transfer, six crown clauses, naturality-aware transport where recorded (\Cref{sec:generalized-crown}).
\item \textbf{Upward explanatory necessity.} Later regimes indispensable for generator-structural truths (\Cref{sec:upward-necessity}).
\item \textbf{Orders, retro, contrasts, philosophy.} Explanatory vs.\ causal order, retroactive revelation, non-equivalences, and the third option (\Cref{sec:order-divergence}--\Cref{sec:philosophical}).
\end{enumerate}

\begin{figure}[htbp]
  \centering
  \begin{tikzpicture}[
    font=\small,
    grnode/.style={draw, rounded corners=2pt, inner sep=5pt, align=center},
    ca/.style={-Stealth, thick},
    ex/.style={-Stealth, thick, dashed},
    up/.style={-Stealth, line width=0.88pt, red!72!black}
  ]
    \node[grnode, fill=blue!8, minimum width=2.45cm] (G) at (0,2.35)
      {\textbf{Lawful generator}\\[-2pt]\footnotesize (fixed update)};
    \node[grnode] (P0) at (3.15,1.15) {$P_0$\\[-2pt]\footnotesize (phase)};
    \node[grnode] (P1) at (3.15,2.35) {$P_1$\\[-2pt]\footnotesize (phase)};
    \node[grnode] (P2) at (3.15,3.55) {$P_2$\\[-2pt]\footnotesize (phase)};
    \node at (3.15,4.45) {$\vdots$};
    \node[grnode, fill=green!7] (R0) at (6.55,1.15) {Regime $R_0$};
    \node[grnode, fill=green!7] (R1) at (6.55,2.35) {Regime $R_1$};
    \node[grnode, fill=green!7] (R2) at (6.55,3.55) {Regime $R_2$};
    \node at (6.55,4.45) {$\vdots$};

    \draw[ca] (G) -- (P0)
      node[pos=0.5, below=5pt, font=\scriptsize, inner sep=1pt] {causal};
    \draw[ca] (P0) -- (P1);
    \draw[ca] (P1) -- (P2);

    \draw[ex] ($(P0.east)-(0.06,0)$) -- ($(R0.west)+(0.06,0)$);
    \draw[ex] ($(P1.east)-(0.06,0)$) -- ($(R1.west)+(0.06,0)$);
    \draw[ex] ($(P2.east)-(0.06,0)$) -- ($(R2.west)+(0.06,0)$);

    \node[grnode, fill=orange!12, text width=2.9cm, anchor=north west] (Efail) at (9.05,4.62)
      {\footnotesize Every fixed admissible reducer $E$ fails somewhere on the tower};

    % Upward necessity: arc stays in a high band (well above $P_2$ and $\vdots$)
    \draw[up]
      (R2.west)
      .. controls ($(R2.west)+(-0.75,2.55)$) and ($(G.north east)+(0.05,1.9)$) .. (G.north east)
      node[pos=0.37, font=\scriptsize, text=red!85!black, align=center, inner sep=2pt,
           fill=white, fill opacity=0.94, text opacity=1, rounded corners=1.5pt] {upward\\necessity};

    % Representative feedback: east of $R_1$, then \emph{past} $R_0$ on the far right, then down and under $P_0$
    \coordinate (r1Exit) at ($(R1.east)+(0.28,0)$);
    \coordinate (pitLow) at ($(P0.south)+(0,-0.52)$);
    \coordinate (genApproach) at ($(G.south)+(-0.1,-0.55)$);
    % Vertical lane cleared to the right of every regime box (same $x$ stack as $R_0$)
    \coordinate (feederEast) at ($(R0.east)+(1.2,0)$);
    \coordinate (feederTop) at (feederEast |- r1Exit);
    \coordinate (feederBot) at (feederEast |- pitLow);
    \draw[up, red!58!black, line width=0.7pt, dashed]
      (R1.east) -- (r1Exit) -- (feederTop) -- (feederBot)
      -- ($(pitLow)+(-1.05,0)$)
      .. controls ($(G.south west)+(-0.55,-0.45)$) and (genApproach) .. (G.south);

  \end{tikzpicture}
  \caption{Conceptual spine: a fixed law drives phases upward in causal order; explanatory regimes attach to phases; every fixed admissible reducer fails somewhere on the tower; upward explanatory necessity relates later regimes to structural truths of the source. Broad transport across aligned natural carriers and organization-scale obstructions (as in \Cref{sec:generalized-crown} and \Cref{tab:strengthened-lean}) show the phenomenon is not encoding-local. Explanatory dependence need not mirror causal order.}
  \label{fig:conceptual-spine}
\end{figure}

What would be unsurprising is a world in which outputs merely outstrip a fixed vocabulary, observers run out of time, or simulation remains expensive. The proved models go further: a single finitely specified generator can causally produce an infinite tower of phases and explanatory regimes such that no fixed admissible explanatory interface is final for that tower, and later generated regimes can become necessary for structural truths about the generator itself.

\begin{keybox}{Six central claims}
\textbf{1. Fixed law without a final explanatory stance.} A single finitely specified generator can produce an infinite family of phases while no fixed admissible explanatory regime uniformly captures them all.

\medskip
\textbf{2. Infinite conservative paradigm towers.} Later regimes can preserve prior phase adequacy and still force genuinely new explanatory structure.

\medskip
\textbf{3. Upward explanatory necessity.} Later generated regimes can become necessary to express or prove structural truths about the generator's own long-run behavior.

\medskip
\textbf{4. Lawful self-transcendence with generalized transfer.} The architecture is neither anti-determinist nor merely rhetorical `emergence'; the flagship pattern survives transport beyond a single distinguished witness model.

\medskip
\textbf{5. Generalized transport under explicit natural hypotheses.} Beyond the witness-level crown, the same strict-gap architecture extends across aligned admissible carriers and supports bounded-organization obstruction and anti-closure results under the stated naturality predicates.

\medskip
\textbf{6. Sharp boundaries at the formal layer.} Conservativity in history is not redundant in the packaged paradigm-shift definition; numeric unbounded recurrence is a genuine load-bearing hypothesis, equivalent at that layer to non-uniform boundedness of the trace; and no blanket ``universal escalation'' of recurrence across \emph{all} lawful numeric generators is claimed or holds (\Cref{sec:boundaries}, \Cref{tab:theorem-inventory}).
\end{keybox}

\noindent
Expository order follows Summits I--VII above and the table of contents. \Cref{sec:premises-novelty-crown} is the entry point for mapping informal worries to formal quantifiers; proof sketches and certificate chains begin in \Cref{sec:witnesses-proofs}.

\section{What would be trivial, and why this is not that}\label{sec:nottrivial}

The theory would be uninteresting if it merely proved one of the following:

\begin{itemize}[leftmargin=2em]
\item a fixed language eventually fails on sufficiently rich outputs;
\item complexity can outrun a bounded observer;
\item prediction can become hard or impossible;
\item a process can have infinitely many outputs;
\item one can always invent richer descriptive vocabulary after the fact.
\end{itemize}

Each of these is real, but none is enough. None produces the kind of inversion the crown construction targets.

The actual target is a much stronger conjunction: a finitely specified generator $G$ and a family of phases and regimes $(P_n,R_n)$ such that

\begin{enumerate}[leftmargin=2em]
\item every $P_n$ is generated by $G$;
\item each $R_n$ adequately explains $P_n$ in the formal sense of the theory;
\item each successor transition is conservative over prior history and irreducible to the prior regime;
\item every fixed admissible explanatory interface fails somewhere on the tower;
\item later regimes can become necessary for structural truths about the generator itself.
\end{enumerate}

It is not a theorem of mere output richness, brute infinitude, or observational unpredictability alone. It is a theorem of \emph{non-final explanation under fixed law}.

This is the sense in which the framework is a theory of lawful self-transcendence. The center of gravity is not ``more happens than finite observers compress easily.'' It is: ``fixed lawful generation does not guarantee a final explanatory stance.'' A single fixed generator can causally produce an infinite tower of phases and regimes such that none are uniformly collapsible into a fixed explanatory interface of the relevant kind, and later regimes can become necessary for truths about the generator itself.

\section{Causal generation versus explanatory reducibility}\label{sec:causal-vs-explanatory}

The core distinction is between two relations:

\begin{itemize}[leftmargin=2em]
\item \textbf{Causal generation:} the system produces a phase, state, or phenomenon as part of its lawful trace.
\item \textbf{Explanatory reducibility:} a fixed admissible regime suffices to preserve the relevant structural distinctions, adequacy conditions, and invariant content.
\end{itemize}

These are not the same relation---and treating them as identical is exactly how the hidden creed smuggles in final closure.

To say that a phase is generated is a constructive or causal claim. To say that it is reducible to a fixed admissible regime is stronger: it requires structural adequacy without loss of the certified distinctions the theory cares about.

\begin{definition}[Informal self-transcendence schema]
A generator $G$ is self-transcending, in the core sense used here, if there exists a sequence $(P_n,R_n)_{n\in\mathbb{N}}$ such that:
\[
\forall n,\; \GeneratedBy(G,P_n),
\quad
\forall n,\; \Adeq(R_n,P_n),
\quad
\forall n,\; \ParadigmShift(R_n,R_{n+1};P_{\le n}),
\]
and
\[
\forall E \in \Admissible,\; \exists n,\; \neg \Expl(E,G,P_n,R_n).
\]
\end{definition}

The force of this definition is easy to miss. The theorem is not merely that later phases need later words. It is that \emph{every fixed admissible explanatory reducer fails somewhere}. This is the structural anti-closure statement.

\begin{quote}
\textbf{Causal sufficiency does not imply explanatory closure.}
\end{quote}

That sentence is the lens through which everything below should be read.

\section{Self-transcending generators: a new species of object}\label{sec:self-transcending}

The self-transcending generator is not a rhetorical flourish; it is the field-defining object of the theory. If such objects exist cleanly under honest admissibility hypotheses, the program is worth doing---not as cataloguing a formal artifact, but because they carve out a new structural possibility in the conceptual space of law, explanation, and reduction.

\medskip
\noindent
\textbf{Why this is not ``a complicated dynamical system.''}
A generator with intricate outputs, a hard-to-predict future, or a long run of surprising behavior may still admit a fixed explanatory regime that is final relative to the admissible class. The self-transcending generator is characterized by the \emph{failure} of that finitude: the causal trace is fully owned by one fixed law, while explanatory closure is systematically defeated for every admissible reducer at some height on the tower. Complexity alone does not imply that joint profile.

\medskip
\noindent
\textbf{Why the object is field-defining.}
Much of science and philosophy silently assumes that tightening law and tightening explanation march together. The self-transcending generator is the minimal schematic shape of their divergence: one fixed engine; inexhaustible stratification of adequate regimes; conservative, irreducible transitions; universal failure of any fixed final interface; and (at the crown) upward necessity for truths about the source itself. Theorems about this shape are not marginal remarks about pathological examples; they are theorems about what \emph{lawful explanation} can and cannot promise.

\medskip
\noindent
\textbf{Why existence is philosophically radical.}
If even one such generator fits the formal package, the hidden creed cannot be treated as analytic. It becomes an empirical-conceptual hostage: perhaps nature never realizes the package, but \emph{in principle} fixed law without final explanation is coherent, structured, and provable. That shifts the default: final closure must be earned, not assumed from the fact of generation.

\begin{enumerate}[leftmargin=2em]
\item the phases are generated by the same fixed law;
\item the corresponding regimes are genuinely adequate for the phases they explain;
\item the transitions between regimes are conservative over prior history;
\item those transitions are structurally irreducible;
\item no fixed admissible explanatory interface is final for the whole tower.
\end{enumerate}

\begin{quote}
\textbf{A system can generate what it cannot finitely finish explaining.}
\end{quote}

That slogan marks the place where ordinary complexity talk ends and a new mathematical possibility begins: not merely many states, but lawful generation without fixed explanatory closure.

What makes the self-transcending generator a new mathematical object is therefore not that it has many states, nor that it surprises bounded observers, but that it occupies a new structural category: exact generation without final explanatory closure. If such objects exist, then explanatory non-finality is not a symptom of epistemic weakness alone; it is a lawful structural possibility.

\section{Paradigm shift as a formal object}\label{sec:paradigm-shift}

One of the deepest aims is to rescue ``paradigm shift'' from sociological fog and give it a structural backbone. The target is not arbitrary language change, nor a destructive rewrite of the past, nor metaphorical ``revolution.'' It is a disciplined pattern: non-destructive transitions that nonetheless force genuine novelty in what can be explained and certified.

\subsection{Conservative extension}

The successor regime $R_{n+1}$ must respect what $R_n$ already secured on the available history $H_n$. Conservativity is the guarantee that the revolution is not vandalism: prior adequacy on certified past structure is preserved when passing upward. Without that constraint, the notion collapses into mere replacement; with it, novelty must squeeze through a narrow gate---it must arrive \emph{despite} honest retention of what came before.

\subsection{Irreducibility}

At the same time, $R_{n+1}$ must not be a notational variant of $R_n$ for the new work at stake. There must be generated structure that $R_n$ cannot adequately treat while $R_{n+1}$ can. Irreducibility is the mathematical content of ``genuinely new explanatory structure'': not convenience, but loss of collapse into the prior explanatory basis for the relevant claims.

\subsection{Why this is paradigm shift, not metaphor}

This is why the theorem package around infinite conservative paradigm towers is not a decorative extension of the theory but one of its deepest results. It shows that explanatory revolution need not be either destructive or merely sociological. It can be historically conservative, mathematically necessary, and still irreducible to any fixed earlier admissible regime. In that sense, paradigm shift becomes a structural theorem rather than a metaphor.

\medskip
\noindent
Formally, a paradigm shift here is a transition from $R_n$ to $R_{n+1}$ over $H_n$ such that:
\begin{enumerate}[leftmargin=2em]
\item $R_n < R_{n+1}$;
\item $\ConservativeOver(H_n,R_n,R_{n+1})$;
\item there exists a new generated phase $P_{n+1}$ with
\[
\neg \Adeq(R_n,P_{n+1}) \wedge \Adeq(R_{n+1},P_{n+1});
\]
\item the new regime is not reducible to the old one with respect to the relevant structural content.
\end{enumerate}

\begin{remark}[Minimality of conservativity in the packaged definition]
In the library, the sealed \texttt{ParadigmShift} record is equivalent to a \emph{weak} paradigm step together with
\texttt{ConservativeOver} on the same history (\leanref{NoveltyTheory.Core.MinimalHypotheses.paradigmShift\_iff\_weak\_and\_conservative}).
There are regimes and histories with the weak conjuncts but \emph{without} conservativity, hence without a packaged shift
(\leanref{NoveltyTheory.Ridge.ConservativeSeparationCountermodel.countermodel\_without\_history\_conservativity}).
\end{remark}

\section{The first summit: fixed generation without explanatory terminality}\label{sec:first-summit}

The first summit proves that the explanatory tower is never uniformly absorbed by any fixed admissible explanatory interface.

\[
\exists G\,\exists (P_n,R_n)_{n\in\mathbb{N}}\,
\Bigl[
\forall n\,\GeneratedBy(G,P_n)
\wedge
\forall E\in\Admissible\,\exists n\,\neg \Expl(E,G,P_n,R_n)
\Bigr].
\]

The system remains perfectly lawful; the law generates the whole trace. What fails is final explanation---not because forecast is hard within a fixed architecture, but because \emph{no} architecture of the admissible kind is final across the tower.

This is the first point at which the theory reveals itself as a theory of non-final explanation rather than merely a theory of complicated systems. The point is not that some interface happens to fail. It is that every fixed admissible interface fails somewhere. That is the first true rupture with the hidden creed.

\begin{keybox}{The first summit}
\textbf{A fixed finitely specified generator can produce every relevant phase causally, and yet no fixed admissible explanatory regime is final for the whole generated future.}
\end{keybox}

\section{The second summit: no final internal theory}\label{sec:no-final-internal}

The no-final-internal-theory form is stronger than ``some representation wears out.''
\begin{quote}
There is no admissible internal theory of the relevant type that is final for the generated future.
\end{quote}

Finality itself becomes impossible in the specified class---not a local glitch but explanatory anti-closure:
\[
\text{fixed law} \;\not\Rightarrow\; \text{fixed final internal theory}.
\]

This is already a very strong claim. It is still not the crown.

The significance of this theorem is not merely technical. It blocks a familiar escape route: the thought that perhaps there is still, somewhere in the architecture, a final internal standpoint from which the whole future explanatory story can be closed. In the relevant formal setting, there is not. The failure is not local but architectural.

\section{The generalized final crown package}\label{sec:generalized-crown}

The flagship mathematical export is the \textbf{generalized final crown}: a coordinated theorem package in which the central philosophical inversion becomes fully formal. What preceding summit layers establish existentially, the generalized crown stabilizes across disciplined transport and scoped admissible structure. Read as one backbone rather than as isolated peaks, the package includes, in order of conceptual dependence:

\begin{enumerate}[leftmargin=2em]
\item \textbf{Generalized iterated structural ascent.} Strict proof-theoretic ascent for structurally certified sentences along the packaged regimes.
\item \textbf{Generalized no fixed structural stratum.} No fixed structural stratum of the relevant formal class is final for the generated future---an anti-closure claim stronger than a one-step gap.
\item \textbf{Generalized future defeat of terminality.} Would-be terminality-style universal bounds are defeated in a future-directed, packaged sense for the canonical tower.
\item \textbf{Generalized generator-truth family.} Ascent is tied to truths about the generator itself, not to arbitrary propositional noise.
\item \textbf{Generalized class transfer.} The phenomenon is realized not only in one witness but across aligned carriers in a nontrivial admissible class.
\item \textbf{Generalized non-single-model artifact.} The gap pattern is certified \emph{not} to be merely an artifact of one model, one encoding, or one carrier.
\end{enumerate}

Taken together, these clauses do not merely strengthen antecedent crown packaging. They change its status. What first appears as a dramatic existence result is then shown to persist across a disciplined admissible class. That is what turns the phenomenon from a striking witness into the beginning of a theory.

The crown phenomenon would be dramatically weaker if it held only in one hand-picked model. Transfer is what blocks the cheap escape: ``this is a trick of the presentation.'' When the same structural failure class recurs under disciplined transport while traces remain genuinely distinct in the relevant sense, the philosophical reading strengthens: one is not watching a coding accident but a reproducible pattern of lawful non-finality.

\medskip
\noindent
The $\texttt{Bool}{\times}\mathbb{N}$ twin construction remains the canonical \emph{existential} staging ground for \textbf{class transfer} in the summit package, because it supplies two concrete aligned generators with trace-level mismatch. The naturality-aware theorems in \Cref{tab:strengthened-lean} then push the \emph{same} output-enum witness \texttt{ProvesAtCarrier} gap to \textbf{every} stage and every \texttt{NaturalAdmissibleInstance} (including $\mathbb{N}$-observation carriers), and record organization-scale and anti-closure consequences under the predicates named there.

\begin{remark}[Six crown clauses in \lean{}]
Each clause above has a named certificate in \texttt{GeneralizedFinalCrownPackage}:
\begin{itemize}[leftmargin=2em,itemsep=0.55em]
\item \textbf{Iterated structural ascent.}

\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_iterated\_structural\_ascent}

\item \textbf{No fixed structural stratum} (for every bound $B$).

\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_no\_fixed\_structural\_stratum}

\item \textbf{Future defeat of terminality.}

\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_future\_defeat\_of\_terminality}

\item \textbf{Generator-truth family paired with ascent.}

\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_generator\_truth\_family}

\item \textbf{Class transfer.}

\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_class\_transfer}

\item \textbf{Non-single-model artifact.}

\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_not\_model\_artifact}
\end{itemize}
\end{remark}

\begin{remark}[Natural class transport]
Without enlarging quantifiers illegally beyond their formal scopes, the same architecture yields:
\begin{enumerate}[label=(\roman*),leftmargin=2.4em,itemsep=0.65em]
\item \textbf{Weak naturality / adequacy transport.}

\leanref{NoveltyTheory.Foundation.NaturalityFacts.naturality\_adequacy\_transport}

\item \textbf{Output-enum strict gaps on every aligned \texttt{NaturalAdmissibleInstance}.}

\leanref{NoveltyTheory.Ridge.BroadTransfer.naturalClass\_preserves\_structural\_ascent}

\leanref{NoveltyTheory.Ridge.BroadTransfer.broad\_transfer\_of\_generalized\_crown}

\item \textbf{Obstructions for \texttt{NaturalBoundedOrganization}.}

\leanref{NoveltyTheory.Ridge.UnboundedOrganization.natural\_no\_final\_organization\_theorem}

\item \textbf{Packaged anti-closure on \texttt{NaturalNumericSystem}.}

\leanref{NoveltyTheory.Ridge.NaturalAntiClosure.natural\_anti\_closure\_theorem}
\end{enumerate}
Informal ``inevitability'' claims must match these statements' hypotheses.
\end{remark}


\medskip
\noindent
For witness systems, proof sketches, and \lean{} theorem anchors keyed to the crown package above, see \Cref{sec:witnesses-proofs}.

\section{The crown summit: upward explanatory necessity}\label{sec:upward-necessity}

\textit{This section is the central summit of the argument.}

If there is a single theoremic-philosophical center of gravity, it is here. Preceding summits establish that fixed lawful generation need not yield explanatory terminality. This summit goes further: it shows that later generated regimes can become necessary, not merely for later phases, but for structural truths about the generator itself. \textbf{This is the inversion theorem.}

The prior summits establish non-finality: no fixed reducer in the relevant class closes the tower; no final internal theory of the relevant type settles the future story. What follows is the deeper inversion. Later generated explanatory regimes can become \emph{necessary} for structural truths about the generator itself. The causal order remains unchanged; what changes is explanatory dependence.

\medskip
\noindent
\textbf{Why it sounds contradictory.}
If the generator is the source of everything downstream, it is tempting to treat everything ``real'' about the process as already implicit in the source description. On that hearing, the idea that \emph{later} strata are indispensable for understanding the \emph{earlier source} feels like logical nonsense---as if the effect somehow authored the cause. That hearing conflates generation with explanation.

\medskip
\noindent
\textbf{Why it is not contradictory once the relations are separated.}
Generation answers: what states arise, in what order, under which fixed update? Explanation answers: which admissible regime is adequate for which structural truths, and what can be proved or expressed at which depth? Those are different questions. Nothing in the forward causal order is reversed when a later regime is required to \emph{state or prove} what an earlier regime could not. The source still generates the whole trace; the derivative layers remain causally downstream; what changes is where certain truths become visible, expressible, or provable.

\medskip
\noindent
\textbf{Why this flips reductionism rather than merely limiting it.}
A modest limitation story says: ``we lack a single brilliant final theory because we are finite.'' The crown says more. It says the hierarchy of intelligibility need not admit a final admissible stratum \emph{even in principle} relative to the class, and that later strata can be \emph{unavoidable} for truths about the source. That is not a gap in our patience; it is a structural feature of the explanatory relation under lawful generation.

\medskip
The usual picture runs like this:
\begin{enumerate}[leftmargin=2em]
\item the generator is fundamental;
\item the generated layers are derivative;
\item therefore the generated layers cannot be strictly necessary for understanding the generator except as convenience.
\end{enumerate}

The crown says step (3) can fail.

\[
\exists G,\exists(R_n),\exists(\Phi_n)\;
\forall n,\;
\Phi_n \text{ is a structural truth about } G
\;\wedge\;
\ProvesAt(R_{n+1},\Phi_n)
\;\wedge\;
\neg \ProvesAt(R_n,\Phi_n).
\]

Read verbally, the theorem says: for each stage there is a structural truth about the generator that becomes provable only from a later regime. The generator remains the source of the whole tower, but the tower it generates becomes necessary for understanding the generator itself.

The claim is not about later outputs alone. It is about structural truths of the source process: the derivative becomes necessary for understanding the source.

In the formalization, the schematic $\Phi_n$ can be taken concretely from the retro-structural family \texttt{histSeqUpto}\,$n$---a finite diagonal history of trace-equalities $(i,i)$ for $i<n+1$---which is \texttt{IsStructuralSentenceV2} and satisfies $\texttt{ProvesAt}\,(n{+}1)\,\Phi_n \land \neg\texttt{ProvesAt}\,n\,\Phi_n$. The named certificates for this family and for generator-anchored sentences such as \texttt{traceEq}\,$n\,n$ are recorded in \Cref{tab:crown-lean} and \Cref{sec:witnesses-proofs} (retro-structural family, strict gap at successor depth, generator-truth packaging).

\begin{keybox}{The crown inversion}
\textbf{Later generated explanatory regimes can become necessary for structural truths about the generator itself.}
\end{keybox}

\section{Explanatory order and causal order}\label{sec:order-divergence}

This section does three jobs at once.

\medskip
\noindent
\textbf{(1) Formal separation.}
Causal order is the forward structure of lawful production:
\[
s_0 \mapsto s_1 \mapsto s_2 \mapsto \cdots
\]
Explanatory order tracks where structural truths become expressible, adequate, or provable within the theory's fixed explanatory regimes. The two orders can diverge: a fact can be causally ``early'' while remaining explanatorily ``late.''

\medskip
\noindent
\textbf{(2) No backward causation.}
Upward explanatory necessity does not mean later regimes pull levers in the past. Causal sufficiency of the generator and forward dynamics are untouched. The inversion concerns intelligibility and proof-theoretic access, not forces traveling up the time arrow.

\medskip
\noindent
\textbf{(3) Upward dependence without causal reversal.}
Explanatory dependence can run from higher strata back toward truths about the generator \emph{without} reversing causal priority. Think of it as a second coordinate system laid over the same lawful trajectory: the causal coordinate orders production; the explanatory coordinate orders what must be said to honor the structure. They need not coincide.

\begin{quote}
\textbf{Explanatory order need not mirror causal order.}
\end{quote}

This is the opening for a new conception of explanation. Causal order may still run downward or forward from a fixed source, while explanatory dependence climbs upward through generated strata that the source itself necessitates but does not finitely close. That is the formal core of lawful self-transcendence.

\section{Retroactive structural revelation}\label{sec:retro}

Later regimes need not merely forecast the next segment of trace. Once explanatory order splits from causal order, a sharper phenomenon appears: later regimes can \emph{disclose latent structure in earlier history}---truths about $P_{\le n}$ that were not expressible, provable, or structurally visible from $R_n$, yet become available from $R_m$ for $m>n$.

\[
\exists n<m,\exists\Psi\;
\Bigl[
R_m \vdash \Psi(P_{\le n})
\wedge
R_n \not\vdash \Psi(P_{\le n})
\Bigr].
\]

Retroactive structural revelation is therefore not an incidental byproduct of the theory. It is one of its deepest consequences. Explanation grows not only by extending into the future but by reorganizing the intelligibility of the past. Later regimes do not merely forecast what comes next; they can disclose what had already been there without yet being structurally sayable. The certified template sentences \texttt{histSeqUpto}\,$n$ are exactly of this ``initial segment made explicit'' form; see \Cref{tab:crown-lean} and the retro discussion in \Cref{sec:witnesses-proofs}.

\section{Why this is not chaos, not computational irreducibility, not emergence rhetoric, and not mere bookkeeping}\label{sec:comparisons}

\begin{keybox}{Disagreement coordinates: reject these identifications}
The target is \emph{not} chaos, opaque simulation cost, loose emergence talk, or mere bookkeeping. Four distinct failures of identification:
\begin{itemize}[leftmargin=2em]
\item \textbf{Generation $\neq$ explanation.} Lawful production of a phase does not entail its explanatory reducibility to any fixed admissible regime.
\item \textbf{Simulation $\neq$ explanatory reduction.} Stepping the dynamics can outrun explanatory collapse: exact generation does not hand you final understanding in the packaged sense.
\item \textbf{Observational sameness $\neq$ explanatory sameness.} Matching coarse traces does not settle adequacy for structurally certified truths at issue in the theory.
\item \textbf{Complexity growth $\neq$ explanatory non-finality.} ``Getting richer'' in an informal sense does not by itself imply the systematic failure of every fixed final reducer in an admissible class.
\end{itemize}
\end{keybox}

\subsection{Chaos}

Chaos: sensitive dependence and unstable long-term prediction \emph{within} a fixed explanatory architecture. Novelty theory: that architecture of the admissible kind may not be final across the tower at all. Difficulty $\neq$ systematic non-finality.

\subsection{Not computational irreducibility}

Computational irreducibility concerns the absence of substantial shortcuts for computing later states of a process. The present theory concerns something stronger and different: explanatory anti-closure---even for a fully lawful generated trace, no single regime need be explanatorily final in the relevant sense. One may have exact generation, exact simulation, or observational agreement and still lack explanatory reduction in the relevant formal sense. The issue here is not merely that the future is costly to compute; it is that explanation itself need not close under a fixed admissible regime.

\subsection{Not generic emergence rhetoric}

The framework does not rest on loose ``emergence,'' ``levels,'' or ``useful descriptions.'' It rests on admissible reductions, adequacy, conservative shifts, generator-anchored truths, and proof-theoretic ascent---where later structure can be mathematically necessary and not uniformly collapsible to any fixed prior explanatory stance of the relevant class.

\subsection{Not ``merely engineered'' admissibility}

A common informal reply is that the crown depends on \emph{specially chosen} admissibility or coding. Two distinctions block a quick dismissal. First, explicit \textbf{obstruction lemmas} rule out illegitimate universal diagonal maneuvers over raw interfaces: the trust boundary is written into the formal predicates, not hidden. Second, the \textbf{naturality-aware block} (\Cref{sec:generalized-crown}, \Cref{tab:strengthened-lean}) separates \textbf{weak naturality / adequacy transport} from \textbf{existence of a crown package}, carries the output-enum \texttt{ProvesAtCarrier} gap across all aligned \texttt{NaturalAdmissibleInstance} carriers (not only $\texttt{Bool}{\times}\mathbb{N}$), and states organization / anti-closure consequences under \texttt{NaturalBoundedOrganization} and \texttt{NaturalNumericSystem}. None of this asserts a universal over all deterministic systems; each theorem fixes the admissible quantifiers it actually uses.

\subsection{Why this is not just incompleteness in disguise}\label{sec:not-incompleteness}

The theory is incompleteness-adjacent in atmosphere, but it is not merely Gödel reheated. Its central object is not a single formal calculus outrun by truth in the abstract. Its central object is a generated tower of phases and explanatory regimes under a disciplined admissibility class, together with \emph{no final admissible closure} for that tower. The crown result is stronger still: later generated regimes can become necessary for structural truths about the generator itself. That is a theorem of explanatory non-finality under exact generation, not just a rephrasing of formal incompleteness.

\section{Why this is not just genotype/phenotype}

Genotype/phenotype is a one-step manifestation gap. This theory is an unbounded tower with conservative irreducible succession and upward necessity at the crown. The analogy can at best illuminate the foothills. It cannot capture the theoremic heart of the argument, which is a multi-stage, formally disciplined, upwardly necessary architecture of explanatory non-closure.

\section{Lawful self-transcendence as a third philosophical option}\label{sec:philosophical}

\begin{keybox}{Three stances (interpretive)}
\textbf{Naïve reductionism.} Everything generated from below, therefore finally explainable from below.

\medskip
\textbf{Mystification.} If not reducible, then outside law or science.

\medskip
\textbf{Third option.} \textbf{Lawful self-transcending structure}---fully generated, fully lawful, not finitely explanation-closed (relative to the admissible class at issue), with possible upward necessity as in \Cref{sec:upward-necessity}.
\end{keybox}

\medskip
\noindent
\textbf{Why the mathematics matters.}
The framework supplies a rigorous language for separating mere complexity, mere redescription, genuine structural novelty, true explanatory phase shifts, and the \emph{non-collapse} of explanation under exact generation. Without such distinctions, debates about reduction, emergence, and explanation talk past one another. The point is not only that some formal possibility exists; it is that one can \emph{name} and \emph{prove} a coherent profile of lawful non-finality strong enough to break the hidden creed without breaking naturalism.

If this framework is right, then the deepest lesson is not merely that the future may surprise us, nor that science may need new theories. It is that full lawfulness and full generative completeness do not force final explanatory closure. The space of lawful systems is wider than reductionism assumed. That is why lawful self-transcendence is not only a new mathematical object, but a new philosophical possibility.

\medskip
\noindent
A \textbf{naturality-aware} extension fixes weak principles linking adequacy across stages, checks structural sentence families against them, and transports output-enum ascent across the whole \texttt{NaturalAdmissibleInstance} class together with organization and anti-closure consequences on explicit natural predicates. The formal hypotheses are part of the content, not an afterthought.

\section{What is formally proved, and what lies outside scope}\label{sec:boundaries}

\begin{keybox}{Trust boundary (scope at a glance)}
\textbf{Inside scope (formal).} Packaged lawful generators with iterated structural ascent, failure of fixed final admissible closure, class transfer, generator-truth alignment, naturality-aware transport under the predicates named in \Cref{tab:strengthened-lean}.

\medskip
\textbf{Outside scope.} ``Every deterministic system,'' backward causation, universal raw-interface diagonal, infinite paradigm tower for every regime family, universal \texttt{FutureDefeat} over all lawful numeric generators---unless a cited lemma states otherwise (\leanref{NoveltyTheory.Ridge.UniversalUpwardNecessity.upward\_necessity\_universal\_obstruction}).

\medskip
\textbf{Reader discipline.} Treat any informal headline as provisional until it matches a row in \Cref{tab:crown-lean}, \Cref{tab:strengthened-lean}, or \Cref{tab:theorem-inventory}.
\end{keybox}

The paper's mathematical claims are strong but sharply scoped. In the formalized models, there exist lawful generators with iterated structural ascent, failure of explanatory terminality relative to the admissible class, future defeat of terminality in the packaged sense, generator-anchored truth families, class transfer across a nontrivial family of aligned carriers (including concrete $\texttt{Bool}{\times}\mathbb{N}$ witnesses), and non-single-model-artifact results.

A further family of theorems (summarized in \Cref{tab:strengthened-lean}) gives \textbf{broad transfer} of the output-enum carrier gap for all aligned \texttt{NaturalAdmissibleInstance} values, \textbf{bounded natural organization} obstructions for the standard future predicate, and \textbf{natural anti-closure} on \texttt{NaturalNumericSystem}---each under the hypotheses named in the corresponding \lean{} declarations.

What the formal package does not entail is equally important. It does not imply that every deterministic system exhibits the crown phenomenon, that every conceivable explanatory notion fails globally, that every organization must be infinite, or that causation runs backward. It does not imply a universal diagonal over all raw admissible interfaces, nor an infinite paradigm tower for every singleton-within regime family. Those literal universal glosses are false under the formal predicates where explicit obstruction applies.

This boundary is not a retreat. It is the condition of the theory's strength. The scoped theorem family is already sufficient to establish the existence, structure, and transfer of lawful non-final explanation under fixed law; the naturality-aware block adds rigorously scoped claims that undercut a narrow ``presentation trick'' dismissal.

\section{Conclusion}

\begin{keybox}{Big payoffs}
\textbf{1. The tacit creed fails.} Fixed fundamental law does \emph{not} entail a fixed final admissible explanatory stance for everything important about the generated future: there are rigorous models where the causal trace is complete and explanatory closure is not.

\medskip
\textbf{2. A whole tower, not a glitch.} The phenomenon is packaged as non-finality across phases and regimes, conservative yet irreducible paradigm succession, and (under the cited predicates) obstruction of would-be terminal organization---not a vocabulary shortfall or a one-off surprise.

\medskip
\textbf{3. Crown inversion. Later regimes can be unavoidable for structural truths about the source.} Explainability and proof-theoretic access need not track causal priority; upward explanatory dependence is compatible with forward generation.

\medskip
\textbf{4. Transfer, not a single encoding.} Class transfer and the naturality-aware block show the flagship gap pattern and companion obstructions are stable across aligned carriers and explicit natural hypotheses---not a trick of one hand-picked model (\Cref{sec:generalized-crown}, \Cref{tab:strengthened-lean}).

\medskip
\textbf{5. Honest quantifiers + machine proof.} Scope is exactly what the formal statements assert; the formal development cited here is machine-checked end to end (\Cref{tab:crown-lean}--\Cref{tab:theorem-inventory}, \Cref{sec:witnesses-proofs}).

\medskip
\textbf{6. Minimality and numeric sharpness land in the library.} Conservativity can fail while weak paradigm-step conjuncts still hold; numeric \texttt{FutureDefeat} aligns with trace dichotomies and admits neither unrestricted universal generalization nor silent strengthening of hypotheses without new theorems (\Cref{tab:theorem-inventory}).
\end{keybox}

The hidden creed with which the argument began can no longer be treated as self-evident. Fixed law does not imply a final explanatory stance.

In the proved models, a single lawful generator can produce an infinite tower of phases and explanatory regimes while defeating every fixed admissible explanatory interface of the relevant class. The succession can be conservative without being reducible, historically preserving what came before while forcing genuinely new explanatory structure. At the crown, later generated regimes can become necessary for structural truths about the generator itself.

That conjunction is the paper's real theoremic contribution. It is stronger than complexity growth, stronger than mere unpredictability, stronger than loose emergence talk, and stronger than a one-step manifestation gap. It identifies a new structural possibility: exact generation without final explanatory closure.

The philosophical consequence is equally clear. Explanation need not bottom out where causation bottoms out. A system may be fully lawful and fully generative without being finitely explanation-closed.

If that possibility is genuinely formal, then a new mathematics begins here: not merely a mathematics of what fixed laws generate, but a mathematics of non-final explanation under fixed law.

\medskip
\noindent
Under the natural-admissibility hypotheses of \Cref{tab:strengthened-lean}, the witness-level ascent is stable across aligned observation encodings, and organization-scale anti-closure is available on the corresponding predicates---so the crown is not a single-carrier quirk.

\appendix

\section{Witness systems, proof sketches, and \texorpdfstring{\lean{} anchors}{Lean~4 anchors}}\label{sec:witnesses-proofs}

\medskip
\noindent\textbf{Code availability.} Lean~4 library \texttt{novelty-theory-lean} (\path{NoveltyTheory/}). Build: \texttt{lake build}. Summit certificates cited in the main text are checked in that source (see tables below for declarations and \lean{} names).

\medskip
This appendix lists formal witnesses, the principal \lean{} certificates, and compact proof sketches. Gray \textbf{At a glance} blocks summarize each witness; blue \textbf{Proof sketch} boxes isolate one-line arguments where useful.

\subsection{Generative systems and the canonical counter}

\begin{witnessmeta}
\textbf{At a glance.}\quad A discrete-time deterministic generator emits an infinite observable trace. The default dynamical witness is a bare counter on $\mathbb{N}$ whose trace is the identity.
\end{witnessmeta}

\begin{itemize}[leftmargin=1.6em, itemsep=0.35em]
\item \textbf{Formal generator.} State $S$, observations $X$, parameters $(s_0,\tau,\mathrm{out})$, trace $\mathrm{trace}_G(n)=\mathrm{out}(\tau^n(s_0))$.
\item \textbf{\lean{} structure.} \leanref{NoveltyTheory.Core.GenerativeSystem}
\item \textbf{Canonical counter \texttt{natCounter.}} $S=X=\mathbb{N}$, $s_0=0$, $\tau=\texttt{Nat.succ}$, $\mathrm{out}=\mathrm{id}$; hence every $n$ is observed at time $n$.
  \begin{itemize}[leftmargin=1.2em, nosep, topsep=2pt]
  \item Definition: \leanref{NoveltyTheory.Models.SignatureTower.natCounter}
  \item Trace: \leanref{NoveltyTheory.Models.SignatureTower.natCounter\_trace}
  \item Singleton phases \texttt{phaseSingleton $k$} are generated: \leanref{NoveltyTheory.Models.SignatureTower.natCounter\_generated}
  \end{itemize}
\end{itemize}

\subsection{Sentences and depth-indexed derivability}

\begin{witnessmeta}
\textbf{At a glance.}\quad Truth \texttt{HoldsAt natCounter} and derivability \texttt{ProvesAt $m$} are defined by recursion on the shape of a \texttt{Sentence}. The critical gate for crown witnesses is a uniform inequality tying \emph{list content} to \emph{proof depth}.
\end{witnessmeta}

\begin{itemize}[leftmargin=1.6em, itemsep=0.35em]
\item \textbf{Depth calculus.} Base depth calculus: \leanref{NoveltyTheory.Models.InvariantTower.provesAtDepth}. Sentence-level derivability: \leanref{NoveltyTheory.Models.SentenceProvability.ProvesAt}.
\item \textbf{Example clause (\texttt{outputEnumMem}\,$\ell\,x$).} Requires:
  \begin{enumerate}[label=\alph*), leftmargin=1.4em, itemsep=0.2em, topsep=2pt]
  \item $x\in\ell$ (the named output appears in the finite list);
  \item $\forall y\in\ell,\; y\le m$ (no list element exceeds the \emph{current} depth $m$);
  \item a depth-$m$ certificate for the singleton-phase atom at $x$.
  \end{enumerate}
\end{itemize}

\noindent
Raising $m$ is what allows longer lists to become uniformly licensable: later depth unlocks stricter global bounds on $\ell$.

\subsection{The output-enumeration gap (transferable witness)}

\begin{witnessmeta}
\textbf{At a glance.}\quad Use the list $\{0,\ldots,n+1\}$ and ask about membership of $n$. The sentence is \emph{true} on \texttt{natCounter} at every stage, but derivability at depth $n$ fails because $n+1\in\ell$ breaks the ``all entries $\le n$'' side condition. One more step of depth repairs the bound.
\end{witnessmeta}

\noindent\textbf{Definition (paper abbreviation).}\quad
$\textsc{OutEnum}_n \;:=\; \texttt{outputEnumMem (List.range (n+2))}\; n$.

\[
  \texttt{ProvesAt}\,(n+1)\,\textsc{OutEnum}_n \;\wedge\; \neg\texttt{ProvesAt}\,n\,\textsc{OutEnum}_n
\]

\begin{itemize}[leftmargin=1.6em, itemsep=0.3em]
\item \textbf{Main certificate.} \leanref{NoveltyTheory.Models.OutputEnumCrownFamily.outputEnumCrownWitness\_proves\_succ\_not\_at}
\item \textbf{List / bound lemmas.}
  \begin{itemize}[leftmargin=1.4em,nosep,topsep=2pt]
  \item \leanref{NoveltyTheory.Models.OutputEnumCrownFamily.mem\_range\_self\_lt\_add\_two}
  \item \leanref{NoveltyTheory.Models.OutputEnumCrownFamily.forall\_mem\_range\_le\_succ\_self}
  \end{itemize}
\item \textbf{Structural status / non-collapse.}
  \begin{itemize}[leftmargin=1.4em,nosep,topsep=2pt]
  \item \leanref{NoveltyTheory.Models.OutputEnumCrownFamily.isStructural\_outputEnumCrownWitness}
  \item \leanref{NoveltyTheory.Models.OutputEnumCrownFamily.outputEnumMem\_not\_geOutput\_abbrev}
  \end{itemize}
\item \textbf{Carrier-general gap (any aligned section).} For an arbitrary observation type $X$ and \texttt{CarrierSection $X$}, the embedded output-enum witness has the same strict \texttt{ProvesAtCarrier} profile: \leanref{NoveltyTheory.Foundation.PhaseGeneralizationFacts.carrier\_outputEnum\_gap\_at}.
\item \textbf{Broad natural class.} For every \texttt{NaturalAdmissibleInstance} (definitionally: aligned generator + section + trace alignment), the same pattern holds at every stage. Packaged conjunction and local lemma:
  \begin{itemize}[leftmargin=1.4em,nosep,topsep=2pt]
  \item \leanref{NoveltyTheory.Ridge.BroadTransfer.broad\_transfer\_of\_generalized\_crown}
  \item \leanref{NoveltyTheory.Ridge.BroadTransfer.naturalClass\_preserves\_structural\_ascent}
  \end{itemize}
\end{itemize}

\begin{proofsketch}
\begin{enumerate}[leftmargin=1.4em, itemsep=0.35em]
\item \textbf{At $m=n+1$.} Check $n\in\texttt{range}(n+2)$, verify $\forall y\in\ell,\; y\le n+1$, supply the phase atom at $n$ (uses \texttt{upward\_phase\_derivability\_gap}).
\item \textbf{At $m=n$.} Assume a proof; instantiate the uniform bound at $y=n+1\in\ell$ to force $n+1\le n$, contradiction.
\end{enumerate}
\end{proofsketch}

\subsection{Retro-structural ascent (\texttt{histSeqUpto})}

\begin{witnessmeta}
\textbf{At a glance.}\quad A second crown family bundles many earlier diagonal trace-equalities $(i,i)$ for $i<n{+}1$. It is certified retro-structural (\texttt{IsStructuralSentenceV2}) and satisfies the same strict $\texttt{ProvesAt}$ gap by a diagonal counting argument at depth.
\end{witnessmeta}

\begin{itemize}[leftmargin=1.6em, itemsep=0.35em]
\item \textbf{Definition.} \leanref{NoveltyTheory.Foundation.RetroStructuralTruthV2.histSeqUpto}
\item \textbf{Strict ascent.} \leanref{NoveltyTheory.Ridge.RetroStructuralGap.histSeqUpto\_proves\_succ\_not\_at}
\item \textbf{Uniform family packaging.} \leanref{NoveltyTheory.Foundation.StructuralSentenceHierarchyV2.structuralV2\_crown\_family}
\item \textbf{Summit re-export.} \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_iterated\_structural\_ascent}
\end{itemize}

\begin{proofsketch}
At $n{+}1$, every pair in the finite \texttt{histSeq} list has time index $<n{+}1$, so each diagonal atom is an eligible fringe proof. At $n$, the pair $(n,n)$ is still in the list but the diagonal trace atom \texttt{traceEq $n\,n$} is not derivable at depth $n$ (\leanref{NoveltyTheory.Models.InvariantTower.not\_proves\_trace\_diag}).
\end{proofsketch}

\subsection{Future defeat of terminality}

\begin{witnessmeta}
\textbf{At a glance.}\quad \texttt{FutureDefeat $G$}: for every proposed output bound, some later time exceeds it. For \texttt{natCounter}, take $t=\texttt{bound}+1$.
\end{witnessmeta}

\begin{itemize}[leftmargin=1.6em, itemsep=0.3em]
\item \textbf{Predicate.} \leanref{NoveltyTheory.Models.StratifiedFinality.FutureDefeat}
\item \textbf{Closed proof.}
  \begin{itemize}[leftmargin=1.4em,nosep,topsep=2pt]
  \item \leanref{NoveltyTheory.Models.StratifiedFinality.natCounter\_futureDefeat}
  \item Uses trace: \leanref{NoveltyTheory.Models.SignatureTower.natCounter\_trace}
  \end{itemize}
\item \textbf{Sharp numeric boundary (no extra crown assumptions).}\quad
  On observation type $\mathbb{N}$, failure of \texttt{FutureDefeat} is equivalent to a uniformly bounded numeric trace; dichotomy:
  \leanref{NoveltyTheory.Ridge.ClosureCollapseBoundary.futureDefeat\_or\_existsBoundedNumericTrace}.
\item \textbf{Not universal over all generators.}\quad
  Constant observable output obstructs \texttt{FutureDefeat} for a lawful system on $\mathbb{N}$; hence
  \leanref{NoveltyTheory.Ridge.UniversalUpwardNecessity.upward\_necessity\_universal\_obstruction}.
\end{itemize}

\subsection{Paradigm packaging: conservativity is not decorative}

\begin{witnessmeta}
\textbf{At a glance.}\quad A \emph{packaged} paradigm shift layers history conservativity on top of a strictly weaker ``weak step'' record. Dropping conservativity can leave all three weak conjuncts while still failing the full package.
\end{witnessmeta}

\begin{itemize}[leftmargin=1.6em, itemsep=0.3em]
\item \textbf{Equivalence.} \leanref{NoveltyTheory.Core.MinimalHypotheses.paradigmShift\_iff\_weak\_and\_conservative}
\item \textbf{Separation witness.} \leanref{NoveltyTheory.Ridge.ConservativeSeparationCountermodel.countermodel\_without\_history\_conservativity}
\end{itemize}

\begin{remark}
The witness uses singleton phases on $\mathbb{N}$ and differing output-set adequacy predicates; it is meant to pin a \emph{definition-theoretic} point, not to model physical dynamics.
\end{remark}

\subsection{No fixed structural stratum (organization height)}

\begin{witnessmeta}
\textbf{At a glance.}\quad Fix a finite organizational height $B$. Some \texttt{natCounter}-generated phase is invisible to the corresponding finite signature stratum, and no abstract height-$B$ organization can \emph{totalize} the standard future ordering on stages.
\end{witnessmeta}

\begin{itemize}[leftmargin=1.6em, itemsep=0.3em]
\item \textbf{Packaged obstruction.} \leanref{NoveltyTheory.Ridge.OrganizationAbstractObstruction.organization\_obstruction\_supports\_generalized\_crown}
\item \textbf{Ingredients (for manual tracing).}
  \leanref{NoveltyTheory.Models.SignatureTower.finite\_signature\_cannot\_organize\_full\_ladder},\;
  \leanref{NoveltyTheory.Ridge.OrganizationHeightObstruction.no\_finite\_adequate\_organization\_totalizes\_future}
\end{itemize}

\subsection{\texorpdfstring{$\texttt{Bool}{\times}\mathbb{N}$ carriers, alignment, and class transfer}{Bool x N carriers, alignment, and class transfer}}

\begin{witnessmeta}
\textbf{At a glance.}\quad Twin laws disagree on a \texttt{Bool} tag at every time, yet each twins the counter after embedding. The same output-enumeration gap survives on the carrier calculus (\texttt{ProvesAtCarrier}). This symmetric pair is the canonical \emph{existential} witness for \textbf{class transfer} in \texttt{GeneralizedFinalCrownPackage}; broad transfer across \texttt{NaturalAdmissibleInstance} is the natural-class extension recorded below.
\end{witnessmeta}

\begin{tabular}{@{}p{0.22\linewidth}p{0.72\linewidth}@{}}
\toprule
\textbf{Object} & \textbf{Role / reference} \\
\midrule
\texttt{trueBoolProdNat} &
Always-\texttt{true} tag; trace $n\mapsto(\texttt{true},n)$. \leanref{NoveltyTheory.Ridge.CrownTransfer.trueBoolProdNat} \\
\texttt{altBoolProdNat} &
Always-\texttt{false} tag; same $\mathbb{N}$ skeleton. \leanref{NoveltyTheory.Ridge.CrownTransfer.altBoolProdNat} \\
\texttt{trace\_true\_vs\_alt} &
Pointwise inequality on traces. \leanref{NoveltyTheory.Ridge.CrownTransfer.trace\_true\_vs\_alt} \\
Alignment &
\texttt{true\_bool\_aligns}, \texttt{alt\_bool\_aligns}.\newline
\leanref{NoveltyTheory.Ridge.CrownTransfer.true\_bool\_aligns}\newline
\leanref{NoveltyTheory.Ridge.CrownTransfer.alt\_bool\_aligns} \\
Faithful embed &
Pull \texttt{natCounter} truth back and forth. \leanref{NoveltyTheory.Foundation.PhaseGeneralizationFacts.phase\_general\_embed\_faithful} \\
Carrier gap &
\leanref{NoveltyTheory.Ridge.CrownTransfer.true\_general\_phase\_gap\_at} (and symmetric \texttt{alt\_\ldots}) \\
Class transfer &
\leanref{NoveltyTheory.Ridge.CrownTransfer.crown\_transfers\_across\_admissible\_class} \\
Non-artifact &
\leanref{NoveltyTheory.Ridge.CrownTransfer.generalized\_crown\_not\_single\_model\_artifact} \\
\bottomrule
\end{tabular}

\begin{proofsketch}
\texttt{true\_general\_phase\_gap\_at}: forward direction refines a purported depth-$n$ carrier proof of the embedded witness to a depth-$n$ proof of the underlying \texttt{Sentence $\mathbb{N}$}, contradicting \texttt{outputEnumCrownWitness\_proves\_succ\_not\_at}. Transfer bundles that gap with an existential witness for \texttt{G} and aligned \texttt{C}; non-artifact instantiates two concrete aligned generators and the first time mismatch.
\end{proofsketch}

\subsection{Natural admissible class, broad transfer, and twin encodings}

\begin{witnessmeta}
\textbf{At a glance.}\quad \texttt{NaturalAdmissibleInstance} packages any observation type $\alpha$ with a generative system, a \texttt{CarrierSection $\alpha$}, and alignment between numeric trace and embedded observations. \texttt{carrier\_outputEnum\_gap\_at} implies \texttt{naturalClass\_preserves\_structural\_ascent} and the staged conjunction \texttt{BroadTransferStatement}. Two $\texttt{Bool}{\times}\mathbb{N}$ instances with differing traces show that the gap is not an encoding artifact \emph{within} that class.
\end{witnessmeta}

\begin{itemize}[leftmargin=1.6em, itemsep=0.3em]
\item \textbf{Aligned instance.} \leanref{NoveltyTheory.Core.NaturalAdmissibleClass.NaturalAdmissibleInstance}
\item \textbf{Concrete shapes.}
  \begin{itemize}[leftmargin=1.4em,nosep,topsep=2pt]
  \item \leanref{NoveltyTheory.Models.NaturalClassExamples.natCounterAdmissible}
  \item \leanref{NoveltyTheory.Models.NaturalClassExamples.boolTrueAdmissible}
  \item \leanref{NoveltyTheory.Models.NaturalClassExamples.boolFalseAdmissible}
  \end{itemize}
\item \textbf{Broad transfer.} \leanref{NoveltyTheory.Ridge.BroadTransfer.broad\_transfer\_of\_generalized\_crown}
\item \textbf{Trace separation + non-artifact.}
  \begin{itemize}[leftmargin=1.4em,nosep,topsep=2pt]
  \item \leanref{NoveltyTheory.Ridge.BroadTransfer.multiple\_nontrivially\_distinct\_instances}
  \item \leanref{NoveltyTheory.Ridge.BroadTransfer.crown\_not\_encoding\_artifact\_in\_natural\_class}
  \end{itemize}
\item \textbf{Organization inevitability layer.} \leanref{NoveltyTheory.Ridge.UnboundedOrganization.natural\_no\_final\_organization\_theorem}
\item \textbf{Anti-closure packaging.} \leanref{NoveltyTheory.Ridge.NaturalAntiClosure.natural\_anti\_closure\_theorem}
\end{itemize}

\subsection{Generator-truth linkage}

\begin{witnessmeta}
\textbf{At a glance.}\quad The summit packages \emph{ascending} structural sentences together with \emph{elementary} generator-structural trace facts \texttt{traceEq $n\,n$} that really hold on \texttt{natCounter}.
\end{witnessmeta}

\begin{itemize}[leftmargin=1.6em, itemsep=0.3em]
\item \textbf{Theorem.} \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_generator\_truth\_family}
\item \textbf{Proof idea.} Split the conjunction: the ascent family is \texttt{structuralV2\_crown\_family}; each \texttt{traceEq $n\,n$} is generator-structural and true by definition of \texttt{natCounter.trace}.
\end{itemize}


\section{Machine-checked certificates and reproducibility}

Every theorem statement cited from \texttt{NoveltyTheory/} is checked by the Lean~4 proof assistant (\textbf{422} \texttt{theorem}/\texttt{lemma} declarations by automated parse of the library sources, \textbf{zero} \texttt{sorry}). Readers can rebuild the library from the public sources using standard \texttt{lake} workflows; the intermediate lemma names in \Cref{tab:theorem-inventory} point into that same tree. The \textbf{107} distinct \texttt{NoveltyTheory.\ldots} paths invoked via \verb|\leanref| in this file resolve to existing modules and declarations (verification run \textbf{3 April 2026}); optional recheck: \path{scripts/verify_paper_leanrefs.py}.

\section{Theorem tables}

\Cref{tab:crown-lean} lists the six summit declarations of the generalized final crown package. \Cref{tab:strengthened-lean} lists identifiers for the naturality-aware extension (broad transfer, distinct encodings within the natural class, organization obstruction, anti-closure packaging). \Cref{tab:theorem-inventory} surveys supporting theorems and definitions in rough dependency order.

\begin{table}[htbp]
  \raggedright
  \footnotesize
  \begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.42\linewidth}>{\raggedright\arraybackslash}p{0.52\linewidth}@{}}
    \toprule
    Crown clause & \lean{} identifier \\
    \midrule
    Iterated structural ascent &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_iterated\_structural\_ascent} \\
    No fixed structural stratum (parameter \texttt{B}) &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_no\_fixed\_structural\_stratum} \\
    Future defeat of terminality &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_future\_defeat\_of\_terminality} \\
    Generator-truth family + ascent &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_generator\_truth\_family} \\
    Class transfer (\texttt{Bool}$\times\mathbb{N}$) &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_class\_transfer} \\
    Not a single-model artifact &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_not\_model\_artifact} \\
    \bottomrule
  \end{tabular}
  \caption{Generalized final crown package: packaged summit theorems.}
  \label{tab:crown-lean}
\end{table}

\begin{table}[htbp]
  \raggedright
  \footnotesize
  \begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.42\linewidth}>{\raggedright\arraybackslash}p{0.52\linewidth}@{}}
    \toprule
    Natural-class clause & \lean{} identifier \\
    \midrule
    Broad transfer at every stage ($\mathbb{N}$ and $\texttt{Bool}{\times}\mathbb{N}$ instances) &
    \leanref{NoveltyTheory.Ridge.BroadTransfer.broad\_transfer\_of\_generalized\_crown} \\
    Non-encoding artifact in natural class &
    \leanref{NoveltyTheory.Ridge.BroadTransfer.crown\_not\_encoding\_artifact\_in\_natural\_class} \\
    Natural anti-closure (packaged) &
    \leanref{NoveltyTheory.Ridge.NaturalAntiClosure.natural\_anti\_closure\_theorem} \\
    \bottomrule
  \end{tabular}
  \caption{Selected identifiers for naturality-aware transport and anti-closure (quantifiers exactly as in the \lean{} statements).}
  \label{tab:strengthened-lean}
\end{table}

{\scriptsize
\setlength{\tabcolsep}{3pt}
\setlength{\extrarowheight}{4pt}
\renewcommand{\arraystretch}{1.24}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{0.09\linewidth}>{\raggedright\arraybackslash}p{0.38\linewidth}>{\raggedright\arraybackslash}p{0.22\linewidth}>{\raggedright\arraybackslash}p{0.27\linewidth}@{}}
  \caption{Extended theorem inventory for the generalized crown mathematical spine (\textbf{40} curated rows; the library currently contains \textbf{422} \texttt{theorem}/\texttt{lemma} declarations in \texttt{NoveltyTheory/} by source parse, so this table is not exhaustive). ``Depends on'' lists immediate named ingredients in the formal proof, not the full transitive closure. Declaration paths in the table and appendix were cross-checked against the repository on \textbf{3 April 2026}.}\label{tab:theorem-inventory} \\
  \toprule
  Layer & \lean{} declaration & Informal content & Immediate dependencies \\
  \midrule
  \addlinespace[0.35em]
  \endfirsthead
  \multicolumn{4}{c}{\tablename\ \thetable{} (continued)} \\
  \toprule
  Layer & \lean{} declaration & Informal content & Immediate dependencies \\
  \midrule
  \addlinespace[0.35em]
  \endhead
  \midrule
  \multicolumn{4}{r}{\emph{Continued on next page}} \\
  \endfoot
  \bottomrule
  \endlastfoot
  Core &
    \leanref{NoveltyTheory.Core.GenerativeSystem} &
    Deterministic generator $(s_0,\tau,\mathrm{out})$ and trace. &
    \texttt{(primitive)} \\
  Models &
    \leanref{NoveltyTheory.Models.SignatureTower.natCounter} &
    Canonical counter: state $\mathbb{N}$, $\mathrm{succ}$, identity observation. &
    \leanref{NoveltyTheory.Core.GenerativeSystem} \\
  Models &
    \leanref{NoveltyTheory.Models.SignatureTower.natCounter\_trace} &
    $\mathrm{trace}(n)=n$. &
    induction on $n$ \\
  Models &
    \leanref{NoveltyTheory.Models.SignatureTower.natCounter\_generated} &
    Singleton phase $\{k\}$ generated by \texttt{natCounter}. &
    \leanref{NoveltyTheory.Models.SignatureTower.natCounter\_trace} \\
  Models &
    \leanref{NoveltyTheory.Models.SentenceProvability.ProvesAt} &
    Depth-$m$ derivability on \texttt{Sentence}. &
    \leanref{NoveltyTheory.Models.InvariantTower.provesAtDepth} \\
  Models &
    \leanref{NoveltyTheory.Models.InvariantTower.not\_proves\_trace\_diag} &
    Diagonal trace atom not derivable at matching depth. &
    \leanref{NoveltyTheory.Models.InvariantTower} \\
  Models &
    \leanref{NoveltyTheory.Models.InvariantTower.upward\_phase\_derivability\_gap} &
    Strict inclusion of derivable facts when depth increases. &
    \leanref{NoveltyTheory.Models.InvariantTower} \\
  Models &
    \leanref{NoveltyTheory.Models.OutputEnumCrownFamily.outputEnumCrownWitness\_proves\_succ\_not\_at} &
    \texttt{outputEnumMem} list witness: provable at $n{+}1$, not at $n$. &
    range lemmas\newline
    \leanref{NoveltyTheory.Models.InvariantTower.upward\_phase\_derivability\_gap} \\
  Models &
    \leanref{NoveltyTheory.Models.StratifiedFinality.natCounter\_futureDefeat} &
    No uniform bound on \texttt{natCounter} trace values (\texttt{FutureDefeat}). &
    \leanref{NoveltyTheory.Models.SignatureTower.natCounter\_trace} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.ClosureCollapseBoundary.not\_futureDefeat\_iff\_existsBoundedNumericTrace} &
    Numeric layer: $\neg$\texttt{FutureDefeat} $\leftrightarrow$ uniformly bounded trace. &
    \leanref{NoveltyTheory.Models.StratifiedFinality.FutureDefeat}\newline
    \leanref{NoveltyTheory.Core.ClosureDichotomy.ExistsBoundedNumericTrace} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.ClosureCollapseBoundary.futureDefeat\_or\_existsBoundedNumericTrace} &
    Dichotomy: unbounded recurrence vs.\ uniform trace bound ($X=\mathbb{N}$). &
    classical case split \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.UniversalUpwardNecessity.upward\_necessity\_universal\_obstruction} &
    $\neg\forall$ lawful $\texttt{GenerativeSystem}$, \texttt{FutureDefeat} (constant trace). &
    \leanref{NoveltyTheory.Ridge.MinimalityCountermodels.not\_futureDefeat\_of\_constant\_numeric\_trace} \\
  Core &
    \leanref{NoveltyTheory.Core.MinimalHypotheses.paradigmShift\_iff\_weak\_and\_conservative} &
    \texttt{ParadigmShift} $\leftrightarrow$ weak step $\land$ \texttt{ConservativeOver}. &
    definitions \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.ConservativeSeparationCountermodel.countermodel\_without\_history\_conservativity} &
    Inhabited weak paradigm step without history conservativity. &
    output-set adequacy on \texttt{phaseSingleton} \\
  Summits &
    \leanref{NoveltyTheory.Summits.ClosureDichotomySummit.closure\_dichotomy\_numeric\_trace} &
    Summit packaging: numeric \texttt{FutureDefeat} $\lor$ bounded-trace disjunction. &
    \leanref{NoveltyTheory.Ridge.ClosureCollapseBoundary.futureDefeat\_or\_existsBoundedNumericTrace} \\
  Summits &
    \leanref{NoveltyTheory.Summits.UniversalUpwardNecessitySummit.upward\_necessity\_universal\_obstruction} &
    Summit re-export: universal \texttt{FutureDefeat} obstruction. &
    \leanref{NoveltyTheory.Ridge.UniversalUpwardNecessity.upward\_necessity\_universal\_obstruction} \\
  Models &
    \leanref{NoveltyTheory.Models.SignatureTower.finite\_signature\_cannot\_organize\_full\_ladder} &
    Fixed signature height misses some singleton phase. &
    \leanref{NoveltyTheory.Models.SignatureTower.tower\_phase\_not\_explained\_by\_fixed\_regime} \\
  Fdn. &
    \leanref{NoveltyTheory.Foundation.RetroStructuralTruthV2.histSeqUpto} &
    Initial-segment diagonal \texttt{histSeq} sentence. &
    \leanref{NoveltyTheory.Core.Sentence} \\
  Fdn. &
    \leanref{NoveltyTheory.Foundation.RetroStructuralTruthV2.isRetroStructuralV2\_histSeqUpto} &
    \texttt{histSeqUpto} is retro-structural \texttt{V2}. &
    list/\texttt{range} facts \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.RetroStructuralGap.histSeqUpto\_proves\_succ\_not\_at} &
    Same strict $\texttt{ProvesAt}$ gap for \texttt{histSeqUpto}. &
    \leanref{NoveltyTheory.Ridge.RetroStructuralGap.histSeqUpto\_proves\_succ}\newline
    \leanref{NoveltyTheory.Ridge.RetroStructuralGap.histSeqUpto\_not\_proves\_at} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.RetroStructuralGap.histSeqUpto\_not\_proves\_at} &
    Failure at $n$: encodes \texttt{histSeqUpto\_mem\_diag} + diagonal obstruction. &
    \leanref{NoveltyTheory.Models.InvariantTower.not\_proves\_trace\_diag} \\
  Fdn. &
    \leanref{NoveltyTheory.Foundation.StructuralSentenceHierarchyV2.structuralV2\_crown\_family} &
    $\Phi_n=\texttt{histSeqUpto}\,n$ with \texttt{IsStructuralSentenceV2} ascent. &
    \leanref{NoveltyTheory.Foundation.RetroStructuralTruthV2.isRetroStructuralV2\_histSeqUpto}\newline
    \leanref{NoveltyTheory.Ridge.RetroStructuralGap.histSeqUpto\_proves\_succ\_not\_at} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.OrganizationHeightObstruction.no\_finite\_adequate\_organization\_totalizes\_future} &
    Finite-height org cannot realize all nested future stage pairs. &
    bound $B$ vs.\ $B{+}1$ on stages \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.OrganizationHeightObstruction.organization\_height\_obstruction} &
    Signature miss + abstract height obstruction (scoped \texttt{OrganizationV2}). &
    \leanref{NoveltyTheory.Models.SignatureTower.finite\_signature\_cannot\_organize\_full\_ladder}\newline
    \leanref{NoveltyTheory.Ridge.OrganizationHeightObstruction.no\_finite\_adequate\_organization\_totalizes\_future} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.OrganizationAbstractObstruction.organization\_obstruction\_supports\_generalized\_crown} &
    Same pattern for \ttwrap{AdmissibleBoundedOrganization}. &
    \leanref{NoveltyTheory.Models.SignatureTower.finite\_signature\_cannot\_organize\_full\_ladder}\newline
    \leanref{NoveltyTheory.Ridge.OrganizationHeightObstruction.no\_finite\_adequate\_organization\_totalizes\_future} \\
  Fdn. &
    \leanref{NoveltyTheory.Foundation.PhaseGeneralizationFacts.phase\_general\_embed\_faithful} &
    \texttt{natCounter} truth $\leftrightarrow$ aligned carrier truth after embed. &
    \leanref{NoveltyTheory.Core.GeneralizedCarrier.AlignsCarrier} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.CrownTransfer.trueBoolProdNat}\newline
    \leanref{NoveltyTheory.Ridge.CrownTransfer.altBoolProdNat} &
    Twin generators on \texttt{Bool}$\times\mathbb{N}$; traces differ on \texttt{Bool}. &
    \leanref{NoveltyTheory.Core.GenerativeSystem} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.CrownTransfer.true\_general\_phase\_gap\_at} &
    Carrier-side \texttt{ProvesAtCarrier} gap for embedded enum witness. &
    \leanref{NoveltyTheory.Models.OutputEnumCrownFamily.outputEnumCrownWitness\_proves\_succ\_not\_at}\newline
    embed equations \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.CrownTransfer.crown\_transfers\_across\_admissible\_class} &
    $\exists$ aligned \texttt{Bool}$\times\mathbb{N}$ system with gap + trace-inequality class-wide. &
    \leanref{NoveltyTheory.Ridge.CrownTransfer.true\_general\_phase\_gap\_at}\newline
    \leanref{NoveltyTheory.Ridge.CrownTransfer.trace\_true\_vs\_alt} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.CrownTransfer.generalized\_crown\_not\_single\_model\_artifact} &
    Two aligned generators with distinct traces. &
    \leanref{NoveltyTheory.Ridge.CrownTransfer.trueBoolProdNat}\newline
    \leanref{NoveltyTheory.Ridge.CrownTransfer.altBoolProdNat}\newline
    alignment lemmas \\
  Summits &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_iterated\_structural\_ascent} &
    Packaged structural \texttt{V2} strict ascent. &
    \leanref{NoveltyTheory.Foundation.StructuralSentenceHierarchyV2.structuralV2\_crown\_family} \\
  Summits &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_no\_fixed\_structural\_stratum} &
    No fixed structural stratum + abstract org obstruction at height $B$. &
    \leanref{NoveltyTheory.Ridge.OrganizationAbstractObstruction.organization\_obstruction\_supports\_generalized\_crown} \\
  Summits &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_future\_defeat\_of\_terminality} &
    \texttt{FutureDefeat} for \texttt{natCounter}. &
    \leanref{NoveltyTheory.Models.StratifiedFinality.natCounter\_futureDefeat} \\
  Summits &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_generator\_truth\_family} &
    \texttt{traceEq} generator truths + structural ascent family. &
    \leanref{NoveltyTheory.Foundation.StructuralSentenceHierarchyV2.structuralV2\_crown\_family}\newline
    trace lemmas \\
  Summits &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_class\_transfer} &
    Transfer of enum crown gap across admissible class. &
    \leanref{NoveltyTheory.Ridge.CrownTransfer.crown\_transfers\_across\_admissible\_class} \\
  Summits &
    \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_not\_model\_artifact} &
    Non-single-model corollary. &
    \leanref{NoveltyTheory.Ridge.CrownTransfer.generalized\_crown\_not\_single\_model\_artifact} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.BroadTransfer.broad\_transfer\_of\_generalized\_crown} &
    Enum witness gap for every natural admissible instance (two observation kinds). &
    \leanref{NoveltyTheory.Foundation.PhaseGeneralizationFacts.carrier\_outputEnum\_gap\_at}\newline
    \leanref{NoveltyTheory.Core.NaturalAdmissibleClass.NaturalAdmissibleInstance} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.BroadTransfer.crown\_not\_encoding\_artifact\_in\_natural\_class} &
    Bool-tag twins + parallel gaps in the natural class. &
    \leanref{NoveltyTheory.Ridge.BroadTransfer.multiple\_nontrivially\_distinct\_instances} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.UnboundedOrganization.natural\_no\_final\_organization\_theorem} &
    No final bounded natural organization certificate for full future. &
    \leanref{NoveltyTheory.Ridge.OrganizationHeightObstruction}\newline
    \leanref{NoveltyTheory.Core.NaturalOrganization} \\
  Ridge &
    \leanref{NoveltyTheory.Ridge.NaturalAntiClosure.natural\_anti\_closure\_theorem} &
    Packaged anti-closure / disjunctive inevitability on natural numeric systems. &
    \leanref{NoveltyTheory.Ridge.NaturalAntiClosure.NaturalNumericSystem}\newline
    weak naturality lemmas \\
\end{longtable}
}

\section{Dependency reference (\texorpdfstring{\lean{} modules and certificate chains}{Lean~4 modules and certificate chains})}

\subsection{Module import closure at the summit}

\texttt{GeneralizedFinalCrownPackage} aggregates the following modules (direct imports), grouped by architectural role:
\begin{description}[style=nextline, font=\normalfont\ttfamily\small]
\item[Core primitives]\mbox{}\\
\begin{itemize}[leftmargin=1.5em,itemsep=0.12em,nosep]
\item \leanref{NoveltyTheory.Core.GenerativeSystem}
\item \leanref{NoveltyTheory.Core.GeneralizedCarrier}
\item \leanref{NoveltyTheory.Core.GeneratorTruth}
\item \leanref{NoveltyTheory.Core.OrganizationAbstract}
\item \leanref{NoveltyTheory.Core.Phase}
\item \leanref{NoveltyTheory.Core.PhaseSyntaxGeneral}
\item \leanref{NoveltyTheory.Core.SentenceSemantics}
\end{itemize}
\item[Foundation]\mbox{}\\
\begin{itemize}[leftmargin=1.5em,itemsep=0.12em,nosep]
\item \leanref{NoveltyTheory.Foundation.PhaseGeneralizationFacts}
\item \leanref{NoveltyTheory.Foundation.StructuralSentenceHierarchyV2}
\end{itemize}
\item[Models]\mbox{}\\
\begin{itemize}[leftmargin=1.5em,itemsep=0.12em,nosep]
\item \leanref{NoveltyTheory.Models.ClassTransferExamples}
\item \leanref{NoveltyTheory.Models.OutputEnumCrownFamily}
\item \leanref{NoveltyTheory.Models.SignatureTower}
\item \leanref{NoveltyTheory.Models.StructuralCrownFamily}
\item \leanref{NoveltyTheory.Models.StratifiedFinality}
\end{itemize}
\item[Ridge]\mbox{}\\
\begin{itemize}[leftmargin=1.5em,itemsep=0.12em,nosep]
\item \leanref{NoveltyTheory.Ridge.CrownTransfer}
\item \leanref{NoveltyTheory.Ridge.OrganizationAbstractObstruction}
\end{itemize}
\item[Summits (prior crown)]\mbox{}\\
\begin{itemize}[leftmargin=1.5em,itemsep=0.12em,nosep]
\item \leanref{NoveltyTheory.Summits.FinalCrownPackage}
\end{itemize}
\end{description}

\noindent\textbf{Boundary certificates (minimality / numeric trace).}
These modules are not in the \texttt{GeneralizedFinalCrownPackage} import closure but supply sharp trust-boundary lemmas cited in the main text: \leanref{NoveltyTheory.Core.MinimalHypotheses}, \leanref{NoveltyTheory.Ridge.ConservativeSeparationCountermodel}, \leanref{NoveltyTheory.Ridge.ClosureCollapseBoundary}, \leanref{NoveltyTheory.Ridge.UniversalUpwardNecessity}.

\noindent\textbf{Imports adjacent to \texttt{GeneralizedFinalCrownPackage} (natural class extension).}
The following modules supply broad transfer and organization / anti-closure certificates alongside the six crown clauses:
\begin{description}[style=nextline, font=\normalfont\ttfamily\small]
\item[Core]\mbox{}\\
\begin{itemize}[leftmargin=1.5em,itemsep=0.12em,nosep]
\item \leanref{NoveltyTheory.Core.NaturalityAxioms}
\item \leanref{NoveltyTheory.Core.StructuralSentenceNaturality}
\item \leanref{NoveltyTheory.Core.NaturalAdmissibleClass}
\item \leanref{NoveltyTheory.Core.NaturalOrganization}
\end{itemize}
\item[Foundation]\mbox{}\\
\begin{itemize}[leftmargin=1.5em,itemsep=0.12em,nosep]
\item \leanref{NoveltyTheory.Foundation.NaturalityFacts}
\item \leanref{NoveltyTheory.Foundation.StructuralSentenceNaturalityFacts}
\end{itemize}
\item[Models]\mbox{}\\
\begin{itemize}[leftmargin=1.5em,itemsep=0.12em,nosep]
\item \leanref{NoveltyTheory.Models.NaturalClassExamples}
\end{itemize}
\item[Ridge]\mbox{}\\
\begin{itemize}[leftmargin=1.5em,itemsep=0.12em,nosep]
\item \leanref{NoveltyTheory.Ridge.AdmissibilityNaturality}
\item \leanref{NoveltyTheory.Ridge.BroadTransfer}
\item \leanref{NoveltyTheory.Ridge.UnboundedOrganization}
\item \leanref{NoveltyTheory.Ridge.NaturalAntiClosure}
\end{itemize}
\item[Summits]\mbox{}\\
\begin{itemize}[leftmargin=1.5em,itemsep=0.12em,nosep]
\item \leanref{NoveltyTheory.Summits.BroadTransferSummit}
\item \leanref{NoveltyTheory.Summits.UnboundedOrganizationSummit}
\item \leanref{NoveltyTheory.Summits.NaturalAntiClosureSummit}
\end{itemize}
\end{description}
\noindent
Transitive imports pull in Mathlib and, among project modules, the following (illustrative, not exhaustive):
\begin{itemize}[leftmargin=2em,itemsep=0.15em,nosep]
\item \leanref{NoveltyTheory.Ridge.OrganizationHeightObstruction}
\item \leanref{NoveltyTheory.Ridge.RetroStructuralGap}
\item \leanref{NoveltyTheory.Models.InvariantTower}
\item \leanref{NoveltyTheory.Foundation.RetroStructuralTruthV2}
\item \leanref{NoveltyTheory.Models.SentenceProvability}
\end{itemize}
The complete \texttt{import} graph is machine-readable; this list is a concise index.

\subsection{Certificate chains (\texorpdfstring{summit $\to$ ingredients}{summit -> ingredients})}

Each named summit theorem below unfolds to the listed \lean{} certificates; indented arrows mark parallel immediate dependencies.

\begin{sloppypar}
\setlength{\emergencystretch}{2.5em}
\begin{itemize}[leftmargin=1.5em, itemsep=1.05em, parsep=0pt]

\item \textbf{Iterated structural ascent.}
\begin{quote}
\footnotesize\raggedright
\setlength{\parskip}{0.4em}
\noindent\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_iterated\_structural\_ascent}\\*
$\hookrightarrow$\;\leanref{NoveltyTheory.Foundation.StructuralSentenceHierarchyV2.structuralV2\_crown\_family}.
\end{quote}

\item \textbf{No fixed structural stratum.}
\begin{quote}
\footnotesize\raggedright
\setlength{\parskip}{0.4em}
\noindent\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_no\_fixed\_structural\_stratum}\\*
$\hookrightarrow$\;\leanref{NoveltyTheory.Ridge.OrganizationAbstractObstruction.organization\_obstruction\_supports\_generalized\_crown}\\*
\hspace*{1.65em}$\hookrightarrow$\;\leanref{NoveltyTheory.Models.SignatureTower.finite\_signature\_cannot\_organize\_full\_ladder}\\*
\hspace*{1.65em}$\hookrightarrow$\;\leanref{NoveltyTheory.Ridge.OrganizationHeightObstruction.no\_finite\_adequate\_organization\_totalizes\_future}.
\end{quote}

\item \textbf{Future defeat of terminality.}
\begin{quote}
\footnotesize\raggedright
\setlength{\parskip}{0.4em}
\noindent\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_future\_defeat\_of\_terminality}\\*
$\hookrightarrow$\;\leanref{NoveltyTheory.Models.StratifiedFinality.natCounter\_futureDefeat}.
\end{quote}

\item \textbf{Generator-truth family.}
\begin{quote}
\footnotesize\raggedright
\setlength{\parskip}{0.4em}
\noindent\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_generator\_truth\_family}\\*
\hspace*{1.65em}$\hookrightarrow$\;\leanref{NoveltyTheory.Foundation.StructuralSentenceHierarchyV2.structuralV2\_crown\_family}\quad\textit{(structural ascent family)}\\*
\hspace*{1.65em}$\hookrightarrow$\;generator-structural \texttt{traceEq} facts (\texttt{isGeneratorStructural\_traceEq}, \texttt{holdsAt\_natCounter\_traceEq\_self}).
\end{quote}

\item \textbf{Class transfer.}
\begin{quote}
\footnotesize\raggedright
\setlength{\parskip}{0.4em}
\noindent\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_class\_transfer}\\*
$\hookrightarrow$\;\leanref{NoveltyTheory.Ridge.CrownTransfer.crown\_transfers\_across\_admissible\_class}\\*
\hspace*{1.65em}$\hookrightarrow$\;\leanref{NoveltyTheory.Ridge.CrownTransfer.true\_general\_phase\_gap\_at} / \texttt{alt\_general\_phase\_gap\_at}\\*
\hspace*{1.65em}$\hookrightarrow$\;\leanref{NoveltyTheory.Models.OutputEnumCrownFamily.outputEnumCrownWitness\_proves\_succ\_not\_at}\\*
\hspace*{1.65em}$\hookrightarrow$\;\textit{carrier embeddings} (phase embed / faithful pullback).
\end{quote}

\item \textbf{Not a single-model artifact.}
\begin{quote}
\footnotesize\raggedright
\setlength{\parskip}{0.4em}
\noindent\leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_not\_model\_artifact}\\*
$\hookrightarrow$\;\leanref{NoveltyTheory.Ridge.CrownTransfer.generalized\_crown\_not\_single\_model\_artifact}\\*
\hspace*{1.65em}\textit{Constructive data:} \texttt{trueBoolProdNat}, \texttt{altBoolProdNat}, alignment lemmas, first-time trace mismatch.
\end{quote}

\item \textbf{Broad transfer across natural admissible instances.}
\begin{quote}
\footnotesize\raggedright
\setlength{\parskip}{0.4em}
\noindent\leanref{NoveltyTheory.Ridge.BroadTransfer.broad\_transfer\_of\_generalized\_crown}\\*
$\hookrightarrow$\;\leanref{NoveltyTheory.Ridge.BroadTransfer.naturalClass\_preserves\_structural\_ascent}\\*
\hspace*{1.65em}$\hookrightarrow$\;\leanref{NoveltyTheory.Foundation.PhaseGeneralizationFacts.carrier\_outputEnum\_gap\_at}\\*
\hspace*{1.65em}$\hookrightarrow$\;\leanref{NoveltyTheory.Core.NaturalAdmissibleClass.NaturalAdmissibleInstance}.
\end{quote}

\item \textbf{Non-encoding artifact in the natural class.}
\begin{quote}
\footnotesize\raggedright
\setlength{\parskip}{0.4em}
\noindent\leanref{NoveltyTheory.Ridge.BroadTransfer.crown\_not\_encoding\_artifact\_in\_natural\_class}\\*
$\hookrightarrow$\;\leanref{NoveltyTheory.Ridge.BroadTransfer.multiple\_nontrivially\_distinct\_instances}.
\end{quote}

\item \textbf{Bounded natural organization obstruction.}
\begin{quote}
\footnotesize\raggedright
\setlength{\parskip}{0.4em}
\noindent\leanref{NoveltyTheory.Ridge.UnboundedOrganization.natural\_no\_final\_organization\_theorem}\\*
$\hookrightarrow$\;\leanref{NoveltyTheory.Ridge.UnboundedOrganization.weakNaturalOrganization\_bounded\_obstruction}\\*
\hspace*{1.65em}$\hookrightarrow$\;\leanref{NoveltyTheory.Ridge.OrganizationAbstractObstruction.no\_bounded\_abstractOrganization\_totalizes\_future}.
\end{quote}

\item \textbf{Natural anti-closure packaging.}
\begin{quote}
\footnotesize\raggedright
\setlength{\parskip}{0.4em}
\noindent\leanref{NoveltyTheory.Ridge.NaturalAntiClosure.natural\_anti\_closure\_theorem}\\*
$\hookrightarrow$\;\leanref{NoveltyTheory.Ridge.NaturalAntiClosure.weak\_natural\_conditions\_imply\_nonfinality\_or\_ascent}.
\end{quote}

\end{itemize}
\end{sloppypar}

\subsection{Conceptual dependency layers (DAG summary)}

\noindent\begingroup
\footnotesize
\setlength{\tabcolsep}{5pt}%
\raggedright
\begin{sloppypar}
\setlength{\emergencystretch}{2.5em}%
\begin{tabular}{@{}>{\centering\arraybackslash}p{4.5em}>{\raggedright\arraybackslash}p{\dimexpr\linewidth-4.5em-\tabcolsep-6pt\relax}@{}}
  \toprule
  \textbf{Height} & \textbf{Role} \\
  \midrule
  0 & \textbf{Ontology}: generative systems, phases, sentences, carrier alignment. \\
  1 & \textbf{Dynamics}: \texttt{natCounter}, trace identities, singleton phases, regimes \texttt{regimeUpto}. \\
  2 & \textbf{Logic at depth}: \texttt{ProvesAt} / \texttt{InvariantTower}, diagonal obstruction \texttt{not\_proves\_trace\_diag}. \\
  3 & \textbf{Witness families}: \texttt{outputEnumCrownWitness}, \texttt{histSeqUpto} strict ascent. \\
  4 & \textbf{Obstructions}: finite signature vs.\ full ladder; finite organization height vs.\ total future. \\
  5 & \textbf{Transfer}: \texttt{Bool}$\times\mathbb{N}$ twins, faithful phase embed, \texttt{ProvesAtCarrier} pullback. \\
  6 & \textbf{Summit}: \texttt{GeneralizedFinalCrownPackage} re-exports. \\
  7 & \textbf{Strengthening}: \texttt{BroadTransfer}, \texttt{UnboundedOrganization}, \texttt{NaturalAntiClosure} + summits. \\
  8 & \textbf{Minimality / numeric boundaries}: packaged paradigm shift vs.\ weak step + \texttt{ConservativeOver}; $\texttt{FutureDefeat}$ vs.\ uniform bounded numeric trace; explicit counterexamples to unrestricted universal glosses. \\
  \bottomrule
\end{tabular}
\end{sloppypar}
\endgroup
\noindent
Heights are expository only: many formal edges cross layers (for example semantic lemmas for \texttt{HoldsAt}).

\section{Obstructions and blocked universals}

Explicit obstruction lemmas refute naive universal diagonal schemes, naive ``infinite paradigm tower for every regime family'' glosses, and the idea that lawful numeric generators must universally realize unbounded observable recurrence. Together with conservativity--separation witnesses and the numeric \texttt{FutureDefeat} equivalence to uniform boundedness (\Cref{tab:theorem-inventory}), these results fix the trust boundary of the admissible predicates. The natural-class theorems in \Cref{tab:strengthened-lean} are \emph{not} blanket universals: each statement carries explicit hypotheses (\texttt{NaturalAdmissibleInstance}, \texttt{NaturalBoundedOrganization}, \texttt{NaturalNumericSystem}, weak naturality). Any informal ``inevitability'' claim must be read against that quantifier structure.

\nocite{*}
\bibliographystyle{plain}
\bibliography{refs}

\end{document}
