
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{0.22\linewidth}>{\raggedright\arraybackslash}p{0.26\linewidth}>{\raggedright\arraybackslash}p{0.22\linewidth}>{\raggedright\arraybackslash}p{0.24\linewidth}@{}}
\toprule
\textbf{Informal claim (draft)} &
\textbf{Lean theorem} &
\textbf{Assumptions / scope} &
\textbf{Interpretation boundary} \\
\midrule
\endfirsthead
\multicolumn{4}{c}{{\tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{Informal claim} & \textbf{Lean theorem} & \textbf{Assumptions} & \textbf{Boundary} \\
\midrule
\endhead
\midrule
\multicolumn{4}{r}{{Continued on next page}} \\
\endfoot
\bottomrule
\endlastfoot

There is a uniform family of sentences that are certified in a
\emph{structural} sense (generator-structural or retro-structural v2)
and exhibit a strict proof-theoretic gap ($\vdash$ at depth $n{+}1$ but
not at $n$) under the packaged calculus.
& \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_iterated\_structural\_ascent}
& Model~C depth calculus (\texttt{ProvesAt}); canonical counter
\texttt{natCounter}; predicates \texttt{IsStructuralSentenceV2}.
& Not a claim about \emph{every} deterministic system or every logic---only
the formalized generative system and sentence universe in the library. \\

No fixed finite explanatory stratum (as formalized via regimes) exhausts
the phenomenon; bounded multiscale organization cannot globally
totalize a standard future diagram of strict $\mathbb{N}$ pairs.
& \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_no\_fixed\_structural\_stratum}
& \texttt{Phase}, \texttt{regimeUpto}, \texttt{explains} packaging;
\texttt{AdmissibleBoundedOrganization}; \texttt{TotalizesStandardFuture} obstruction lemmas.
& Scoped to the formal \texttt{OrganizationV2} / admissible-class
predicates; not a metaphysical claim that ``all organizations'' fail. \\

A formal \emph{future defeat} property holds for the canonical counter
(forbidding a terminality-style universal bound on observable outputs).
& \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_future\_defeat\_of\_terminality}
& \texttt{FutureDefeat} on \texttt{natCounter} as defined in
\texttt{StratifiedFinality}.
& This is a theorem about the \emph{defined} terminality predicate and
the canonical generator, not empirical open-endedness in the world. \\

Generator-truth linkage on the canonical counter coexists with a
structural strict-ascent family (trace-equality atoms $\land$ structural
ascent family in one bundle).
& \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_generator\_truth\_family}
& \texttt{IsGeneratorStructural}, \texttt{HoldsAt natCounter},
\texttt{ProvesAt} gap for the structural-v2 family.
& Bundles two phenomena; does not assert that \emph{every} true-on-counter
sentence is structurally certified. \\

The enum-crown gap pattern is compatible with an admissible \emph{class}
of aligned carriers (two concrete aligned systems on
$\mathbb{N}\to(\mathrm{Bool}\times\mathbb{N})$ with opposing Bool tags).
& \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_class\_transfer}
& \texttt{ProvesAtCarrier}, \texttt{AlignsCarrier}, constructed
\texttt{trueBoolProdNat} / \texttt{altBoolProdNat} pair; enum witness
pullback.
& \textbf{Existential / constructed class:} not ``for all carriers'' or
``for all tags.'' \\

The library exhibits two trace-distinct aligned systems (not a
single-model artifact for the Bool-tagged transfer story).
& \leanref{NoveltyTheory.Summits.GeneralizedFinalCrownPackage.generalized\_final\_crown\_not\_model\_artifact}
& Same transfer construction as previous row; inequality of traces at
some stage.
& Strengthens honesty of the class-transfer headline; still not a
universal across all mathematical generators. \\

\midrule
\multicolumn{4}{@{}p{\linewidth}@{}}{%
\textbf{Global obstructions (boundaries, not headline package).}
There is no uniform diagonal over all raw admissible-interface rows, and no
infinite paradigm-shift chain for every singleton-within regime family---these
are lemma-level refutations of \emph{literal} universal glosses; see the
``Trust boundary'' section in the main text, with representative lemmas such as
\leanref{NoveltyTheory.Ridge.SignatureAdmissibleBundle.trivialInterface\_rows\_claim\_everything}
in the artifact.}
\\
\end{longtable}
